---
title: "Negative binomial regression using nuclei count data, per cell type, for RNA and/or ATAC, invnorm traits- with log(total nuclei) offset."
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, dev=c('png'), include=F}
knitr::opts_chunk$set(dev = c('png'))
knitr::opts_chunk$set(autodep = TRUE)

```


```{r include=F}
.libPaths("/net/snowwhite/home/aujackso/R/x86_64-pc-linux-gnu-library/4.0");
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(broom)
library(MASS)
options(scipen=1, digits=5)
options(nwarnings=10000)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(corrr)
library(gplots) #library with heatmap.2
library(kableExtra)  

invnorm = function(x, seed) {
  set.seed(seed)
  qnorm((rank(x, na.last="keep", ties.method="random") - 0.5) / sum(!is.na(x)))
}

```


```{r functions, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}

run_nb_invnorm<-function(ds0,trt,trt2,cell) {
	print(paste0(trt," ",cell))
	temp<- ds0 %>% dplyr::select(-labelcode) 
	tot.nuc<-rowSums(temp, na.rm=T)
	labelcode<-ds0$labelcode
	temp.totnuc<-cbind.data.frame(labelcode, tot.nuc)
	temp.phen<-merge(temp.totnuc, trait.ds, by="labelcode")
	temp.phen<-dplyr::slice(temp.phen, match(ds0$labelcode, labelcode)) 
	if (trt == "SEX") {
	   phen<-temp.phen %>% dplyr::select(labelcode, age_biopsy, area_name, batch, trt, tot.nuc) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("nuc ~ agescale + BATCH + area_name + x + offset(log(tot.nuc))") 
	}   
	if (trt == "age_biopsy") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, area_name, batch, trt, tot.nuc) %>%
		mutate(BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("nuc ~ SEX + BATCH + area_name + x + offset(log(tot.nuc))") 
	}   
	if (trt == "T2D") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, all_of(trt), tot.nuc) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + x + offset(log(tot.nuc))") 
	}   
	if (!(trt %in% c("SEX", "age_biopsy", "T2D"))) {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc) %>%
		dplyr::filter(complete.cases(.)) %>% 
		dplyr::mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("nuc ~ agescale + SEX + BATCH + area_name + x + offset(log(tot.nuc))") 
	   phen$agescale=scale(phen$age_biopsy)	
	}
	# only use samples with cell type data
	ds<-ds0 %>% dplyr::select(labelcode,paste0(cell)) %>% filter(complete.cases(.))
	names(ds)<-c("labelcode", "nuc")
	
	# ensure count samples are same as complete pheno samples 
	use<-intersect(ds$labelcode, phen$labelcode)
	
	ds<- dplyr::slice(ds, match(use, labelcode)) 
	phen<-dplyr::slice(phen, match(use, labelcode))
	df<- merge(ds, phen, by="labelcode")
	
	nb<-glm.nb(as.formula(fstr), data=df)
	#pois<-glm(as.formula(fstr), family="poisson", data=df)
	#ll<-pchisq(2*(logLik(nb) - logLik(pois)), df=1, lower.tail=FALSE)
	#print(ll)
	res<-tidy(nb) %>% filter(substr(term,1,1)=="x")
	# res is for the last variable in the design
	res$celltype=cell
	res$trait=trt2
	res$n=nrow(df)
	return(res)
}

#adjusting for log(S_Insu)
run_nb_invnorm_ins<-function(ds0,trt,trt2,cell) {
	print(paste0(trt," ",cell))
	temp<- ds0 %>% dplyr::select(-labelcode) 
	tot.nuc<-rowSums(temp, na.rm=T)
	labelcode<-ds0$labelcode
	temp.totnuc<-cbind.data.frame(labelcode, tot.nuc)
	temp.phen<-merge(temp.totnuc, trait.ds, by="labelcode")
	temp.phen<-dplyr::slice(temp.phen, match(ds0$labelcode, labelcode)) %>%
		mutate(lnIns=log(S_Insu))
	if (trt == "SEX") {
	   phen<-temp.phen %>% dplyr::select(labelcode, age_biopsy, area_name, batch, all_of(trt), tot.nuc, lnIns) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("nuc ~ agescale + as.factor(BATCH) + as.factor(area_name) + lnIns + x + offset(log(tot.nuc))") 
	}   
	if (trt == "age_biopsy") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, area_name, batch, all_of(trt), tot.nuc, lnIns) %>%
		mutate(BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("nuc ~ as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + lnIns + x + offset(log(tot.nuc))") 
	}   
	if (trt == "T2D") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, all_of(trt), tot.nuc, lnIns) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + lnIns + x + offset(log(tot.nuc))") 
	}   
	if (!(trt %in% c("SEX", "age_biopsy", "T2D"))) {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, all_of(trt), tot.nuc, lnIns) %>%
		dplyr::filter(complete.cases(.)) %>% 
		dplyr::mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + lnIns + x + offset(log(tot.nuc))") 
	   phen$agescale=scale(phen$age_biopsy)	
	}
	# only use samples with cell type data
	ds<-ds0 %>% dplyr::select(labelcode,paste0(cell)) %>% filter(complete.cases(.))
	names(ds)<-c("labelcode", "nuc")
	
	# ensure count samples are same as complete pheno samples 
	use<-intersect(ds$labelcode, phen$labelcode)
	
	ds<- dplyr::slice(ds, match(use, labelcode)) 
	phen<-dplyr::slice(phen, match(use, labelcode))
	df<- merge(ds, phen, by="labelcode")
	
	nb<-glm.nb(as.formula(fstr), data=df)
	#pois<-glm(as.formula(fstr), family="poisson", data=df)
	#ll<-pchisq(2*(logLik(nb) - logLik(pois)), df=1, lower.tail=FALSE)
	#print(ll)
	res<-tidy(nb) %>% filter(substr(term,1,1)=="x")
	# res is for the last variable in the design
	res$celltype=cell
	res$trait=trt2
	res$n=nrow(df)
	return(res)
}



#dotplot for a single trait, x=celltype
dotplot<- function(ds, trtstr, pthresh) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	ds<-ds %>% 
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

	return(ggplot(ds, aes(x=celltype, y=estimate, color=psig)) + 
	geom_point(size=2) + 
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) + 
  	geom_errorbar(aes(ymin=estimate-1.96*(std.error), ymax=estimate+1.96*(std.error))) + 
	scale_colour_manual(values=myColors)  + 
	geom_hline(yintercept=0, colour="gray") + 
  	ggtitle(paste0(trtstr)) + 
	facet_wrap(~modality, strip.position="left", dir="v")
	)
}



#dotplot for a single celltype, x=traits
dotplot_celltype<-function(ds, cellstr,pthresh,ylim) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	ds<-ds %>% 
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

return(ggplot(ds, aes(x=trait, y=estimate, color=psig)) + 
	geom_point(size=2) + 
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) + 
  	geom_errorbar(aes(ymin=estimate-1.96*(std.error), ymax=estimate+1.96*(std.error))) + 
	scale_colour_manual(values=myColors)  + 
	geom_hline(yintercept=0, colour="gray") + 
  	ggtitle(paste0(cellstr, " NUCLEI")) +
	facet_wrap(~modality, strip.position="left", dir="v")	
	)
}

#dotplots where each facet row is a celltype, x=traits 
dotplot_celltype_rnaatac<-function(ds, types, pthresh, ymin, ymax, sz) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	temp<-ds %>% dplyr::filter(celltype %in% all_of(types))  %>%
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

   return(
	ggplot(temp, aes(x=trait, y=estimate, color=psig)) + 
	geom_point(size=2) + 
	
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5, size=sz)) + 
  	geom_errorbar(aes(ymin=estimate-1.96*(std.error), ymax=estimate+1.96*(std.error))) + 
	ylim(ymin, ymax) + 
	scale_colour_manual(values=myColors)  + 
	theme(legend.position="none") + 
	geom_hline(yintercept=0, colour="gray") + 
	facet_wrap(~celltype, strip.position="right", dir="v")	
   )
}

cor_heatmap<-function(ds, meth, prefixPlot, cex, mrg) {
	# default method=pearson and use=pairwise.complete.obs
	# can also set use to everything, all.obs, complete.obs, na.or.complete
	cor.tibble<-correlate(ds, method=meth)
	cor.mx<-cor.tibble %>% dplyr::select(-term) %>% as.matrix()
	colnames(cor.mx)<-cor.tibble$term
	rownames(cor.mx)<-cor.tibble$term

	return(
	heatmap.2(cor.mx, trace="none", dendrogram="none",
	Rowv=FALSE, Colv=FALSE, 
	cexRow=cex, cexCol=cex, 
	col=bluered(256), margins=c(mrg,mrg),
	offsetRow=-0.5, offsetCol=-0.5, # to move to left try setting to -36
	adjRow=c(0,NA), adjCol=c(NA,0), # Defaults shown; to left justify and center row labels vertically adjRow=c(0,vrtadj0.5), to right justify row labels adjRow=c(1,0.5); to right justify column labels and center horizontally adjCol=c(0,0.5)
	density.info="none", 
	key=TRUE, keysize=0.6, key.par=list(cex=0.3)) 
	)
}




```
## SEX only, RNA+ATAC nuclei, 279 samples  
### Negative binomial model: nuclei counts ~ age.scaled + sex + batch + area_name + offset(log(tot.nuc))  
###  Dot plots are the neg binomial estimates +/- 1.96*SE    
Red/blue 'significant' represents B-H q<.05, calculated across cell types for sex.  

```{r make_negbinom_output_sex, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE, results='asis'}

####  RNA+ATAC  
out<-data.frame()
df.both.nuc<-readRDS("data/df.both.nuc.Rds")
trait.ds<-readRDS("data/trait.ds.281.Rds") #includes 2 extra samples that won't be merged in

traits<-c("SEX")      
traits2<-c("SEX.M") 
  
celltypes<-c("Type_1", "Type_2a", "Type_2x", "Adipocyte", "Endothelial", "Macrophage", "Mesenchymal_Stem_Cell", "Neuromuscular_junction", "Neuronal", "Satellite_Cell", "Smooth_Muscle", "T_cell")

for (j in 1:length(celltypes)) {
   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.both<-run_nb_invnorm(df.both.nuc, traits[i], traits2[i], celltypes[j]) %>% 
		mutate(modality="rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   if (j==1) {out<-tempout} else {out<- rbind.data.frame(out, tempout)}
}
#make B-H adjustment across cell type
out<- out %>% mutate(padj=p.adjust(p.value, method="BH"))

saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_cell_SEX_fdr_across_celltypes.Rds")
out<-out %>% dplyr::select(trait, celltype, n, estimate, std.error, p.value, padj, modality) 
print(kable(out) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F))

```
```{r nb_nuclei_sex, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each celltype for sex"}
out<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_SEX_fdr_across_celltypes.Rds")
print(dotplot(out, "SEX.M", 0.05))

```


###  Repeat for RNA nuclei only  
```{r make_negbinom_output_sex_rna, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE, results='asis'}

####  RNA  
out<-data.frame()
df.rna.nuc<-readRDS("data/df.rna.nuc.Rds")
trait.ds<-readRDS("data/trait.ds.281.Rds") 

traits<-c("SEX")      
traits2<-c("SEX.M") 
  
celltypes<-c("Type_1", "Type_2a", "Type_2x", "Adipocyte", "Endothelial", "Macrophage", "Mesenchymal_Stem_Cell", "Neuromuscular_junction", "Neuronal", "Satellite_Cell", "Smooth_Muscle", "T_cell")

for (j in 1:length(celltypes)) {
   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.both<-run_nb_invnorm(df.rna.nuc, traits[i], traits2[i], celltypes[j]) %>% 
		mutate(modality="rna")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   if (j==1) {out<-tempout} else {out<- rbind.data.frame(out, tempout)}
}
#make B-H adjustment across cell type
out<- out %>% mutate(padj=p.adjust(p.value, method="BH"))

saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_cell_SEX_fdr_across_celltypes_RNAnuc.Rds")
out<-out %>% dplyr::select(trait, celltype, n, estimate, std.error, p.value, padj, modality) 
print(kable(out) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F))

```
```{r nb_nuclei_sex_rna, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each celltype for sex, RNA nuclei only"}
out<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_SEX_fdr_across_celltypes_RNAnuc.Rds") %>% arrange(estimate)
print(dotplot(out, "SEX.M", 0.05))

```


###  Repeat for ATAC nuclei only
```{r make_negbinom_output_sex_atac, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE, results='asis'}

####  ATAC  
out<-data.frame()
df.atac.nuc<-readRDS("data/df.atac.nuc.Rds")
trait.ds<-readRDS("data/trait.ds.281.Rds") 

traits<-c("SEX")      
traits2<-c("SEX.M") 
  
celltypes<-c("Type_1", "Type_2a", "Type_2x", "Adipocyte", "Endothelial", "Macrophage", "Mesenchymal_Stem_Cell", "Neuromuscular_junction", "Neuronal", "Satellite_Cell", "Smooth_Muscle", "T_cell")

for (j in 1:length(celltypes)) {
   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res<-run_nb_invnorm(df.atac.nuc, traits[i], traits2[i], celltypes[j]) %>% 
		mutate(modality="atac")
	if (i==1) {tempout=res} else {tempout=rbind.data.frame(tempout,res)}
   }
   if (j==1) {out<-tempout} else {out<- rbind.data.frame(out, tempout)}
}
#make B-H adjustment across cell type
out<- out %>% mutate(padj=p.adjust(p.value, method="BH"))

saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_cell_SEX_fdr_across_celltypes_ATACnuc.Rds")
out<-out %>% dplyr::select(trait, celltype, n, estimate, std.error, p.value, padj, modality) 
print(kable(out) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F))

```
```{r nb_nuclei_sex_atac, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each celltype for sex, ATAC nuclei only"}
out<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_SEX_fdr_across_celltypes_ATACnuc.Rds") %>% arrange(estimate)
print(dotplot(out, "SEX.M", 0.05))

```







```{r make_negbinom_output_invnorm, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}

out<-data.frame()

df.both.nuc<-readRDS("data/df.both.nuc.Rds")
trait.ds<-readRDS("data/trait.ds.281.Rds") 
# note beware any element in traits starting with "x" since that's the name of the dummy phenotype in the function  
traits<-c("SEX", "age_biopsy", "T2D", "S_Insu", "bmi",  "glu_fast_biopsy", "HEIGHT", "WEIGHT", "WAIST", "whr", "HIP", "S_ALAT", "S_hs_CRP", "fS_Kol", "fS_Kol_HDL", "fS_Kol_LDL_c", "fS_Trigly", "GL30", "GL60", "GL120", "S_LipoA1", "S_LipoB", "fS_C_pept", "fS_Krea", "S_GT", "S_Uraat", "fS_C_pept_30", "S_Insu_30", "B_GHb_A1C", "glu_2h_biopsy", "daily_energy_expenditure", "total_daily_physical_activity", "moderate_physical_activity", "strenuous_physical_activity",  "p_insu", "p_insu_30", "p_insu_60", "p_insu_120", "RFM", "sbp", "dbp", "ApoB_A1_ratio", "Ins_AUC_0to30", "Glu_AUC_0to30", "InsSec30", "InsGenIn", "DIo", "CpepGenIn", "HOMA", "matsuda_4pt", "matsuda_3pt", "LTPA_duration_all", "LTPA_duration_cond", "LTPA_duration_noncond", "LTPA_energy_all", "LTPA_energy_cond", "LTPA_energy_noncond", "LTPA_energy_light", "LTPA_energy_moderate", "LTPA_energy_modvig", "LTPA_energy_vigorous", "sleep_biopsy", "totalcw", "sleep_24h", "GL0")      
traits2<-c("SEX.M", "AGE", "T2D", "INSFAST", "BMI", "GLUFAST", "HEIGHT", "WEIGHT", "WAIST", "WHR", "HIP", "ALAT", "CRP", "CHOL", "HDL", "LDL", "TG", "GL30", "GL60", "GL120", "APOA1", "APOB", "CPEP", "CREAT", "GT", "URIC", "CPEP30", "S_INS30", "HBA1C", "GLU2H", "ENERGY", "PA_TOT", "PA_MOD", "PA_STREN", "INS0", "INS30", "INS60", "INS120", "RFM", "SBP", "DBP", "APOB.A1.RATIO", "INS_AUC_0_30", "GLU_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG", "SLEEP", "ALCOHOL", "SLEEP24H", "GL0")   
 
  
celltypes<-c("Type_1", "Type_2a", "Type_2x", "Adipocyte", "Endothelial", "Macrophage", "Mesenchymal_Stem_Cell", "Neuromuscular_junction", "Neuronal", "Satellite_Cell", "Smooth_Muscle", "T_cell")


for (j in 1:length(celltypes)) {
   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.both<-run_nb_invnorm(df.both.nuc, traits[i], traits2[i], celltypes[j]) %>% 
		mutate(modality="rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   if (j==1) {out<-tempout} else {out<- rbind.data.frame(out, tempout)}
}

saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype.Rds")
```


## All traits
### Negative binomial model: nuclei counts ~ age.scaled + sex + batch + area_name + trait.invnorm  + offset(log(tot.nuc))  
###  Dot plots are the neg binomial estimates +/- 1.96*SE    
Red/blue 'significant' represents B-H q<.05, calculated within each cell type across all 65 traits including sex - so the single trait plots are misleading   

For example, for SEX.M, the rna+atac for Satellite cell is NOT significant when looking across all traits, 
however when calculating p.adj across cell types just for SEX in rna+atac nuclei, SEX is significant in satellite cells.

###  Dot plots for each celltype, per invnorm trait  
```{r nb_nuclei_invnorm, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each celltype, per invnorm trait"}
out<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype.Rds")

for (i in 1:length(traits)) {
tp<-out %>% filter(trait==traits2[i])
print(dotplot(tp, traits2[i], 0.05))
}
```

##  Dot plots for each invnorm trait, per celltype  
```{r celltype_traits_invnorm, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each invnorm trait, per celltype"}


out<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype.Rds")
for (i in 1:length(celltypes)) {
tp<-out[which(out$celltype==celltypes[i]),]
tp$trait<-factor(tp$trait, levels=c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "INSFAST", "INS0", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE"))
print(dotplot_celltype(tp, celltypes[i], 0.05))
}
```


##  Dot plots all traits and celltypes

* SEX(Male): +Type_2x, -Type_1, -Neuronal (as before)
* BMI: +Neuromuscular_junction, +Type_2x, -MFM  (no Type_1 anymore)
* AGE: +Neuromuscular_junction, -Type_2x (Type 2x NEW)
* T2D: +Neuromuscular_junction (NEW)
* Fasting insulin: +Neuromuscular_junction, +Type_2x, -MFM (no Type_1 anymore) 
* Height: none (as before)
* Hip: +Neuromuscular_junction only (previously also had +Type_2x, -MFM)
* Waist: +Neuromuscular_junction, +Type_2x only (previously also had -Type_1, -MFM)
* Weight: +Neuromuscular_junction, +Type_2x only (previously also had -MFM)
* WHR: +Neuromuscular_junction, +Type_2x, -Type_1 only (previously also had  -MFM)
* Fasting glucose: +Neuromuscular_junction (previously only had +Type_2x)
* ALAT: +Neuromuscular_junction, +Type_2x, -Type_1, -MFM (as before)
* CRP: +Neuromuscular_junction (previously also had  +Type_2x)
* Total cholesterol: none (as before)
* HDL: +MFM, -Type_2x (as before)
* LDL: none (as before)
* Triglycerides: +Type_2x, -Type_1, -MFM (as before)
* GL0: +Neuromuscular_junction only (previously also had +Type_2x)
* GL30: +Neuromuscular_junction, +Type_2x (as before)
* GL60: +Neuromuscular_junction, +Type_2x, -MFM (as before)
* GL120: +Neuromuscular_junction, +Type_2x, -MFM (as before)
* GLC_AUC_0_30: +Neuromuscular_junction, +Type_2x (as before)
* GLU2H: +Type_2x, -Type_1, -MFM (previously no - Type_1)
* HbA1c: none (previously +Type_2x)
* ApoA1: +MFM (as before)
* ApoB: none (as before)
* CPEP: +Neuromuscular_junction, +Type_2x, -Type_1, -MFM (as before)
* Creatinine: -Neuromuscular_junction (new)
* GT: +Neuromuscular_junction only (previously also had +Type_2x)
* Uric acid:  +Neuromuscular_junction, +Type_2x (as before)
* CPEP30: +Type_2x, -Type_1, -MFM (as before) 
* INS30_Serum: +Type_2x, -Type_1, -MFM (as before)
* Energy expenditure: +Smooth_muscle (as before)
* PA_TOT: none (previously +MFM)
* PA_MOD: +MFM, +Smooth_muscle (previously only +Smooth_muscle)
* PA_STREN: +MFM (previously +Smooth_muscle)
* ENERGY_ALL: -Type_2x, -Neuromuscular_junction, +Type_1, +MFM (as before)
* ENERGY_COND: -Type_2x, +Type_1, +MFM (previously also had +Smooth_muscle)
* ENERGY_NONCOND: none (as before)
* INS0: +Neuromuscular_junction, +Type_2x, -MFM (as before)
* INS30: +Type_2x (previously also had -MFM)
* INS60: +Type_2x, -Type_1, -MFM (as before)
* INS120: +Type_2x, +Smooth_muscle, -Type_1, -MFM (as before)
* RFM: +Neuromuscular_junction, +Type_2x,  -Type_1, -MFM (as before)
* SBP: +Neuromuscular_junction, +Type_2x (as before)
* DBP: +Neuromuscular_junction (previously also had +Type_2x)
* ApoB/ApoA1: -MFM (as before)
* INS_AUC_0_30: +Type_2x, -Type_1, -MFM (as before)
* INSSEC: +Type_2x, -Type_1, -MFM (as before)
* INSGENIN: +Type_2x (previously -MFM)
* DI: +Type_2x, -Type_1, -MFM (as before)
* CPEPGENIN: none (as before)
* HOMA: +Neuromuscular_junction, +Type_2x, -Type_1, -MFM (as before)
* Matsuda 4pt: -Neuromuscular_junction, -Type_2x, +Type_1, +MFM (previously no -NMJ)
* Matsuda 3pt: -Neuromuscular_junction, -Type_2x, +Type_1, +MFM (as before)
* SLEEP.BIOP: none (as before)
* SLEEP.24H: none (as before)
* Alcohol: +Neuromuscular_junction, -Smooth_muscle (none before)
* Smoker: none (as before)

```{r dot_rna_atac_all, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each invnorm trait for all celltypes"}

types1<-c("Type_1", "Type_2a")
types2<-c("Type_2x", "Neuromuscular_junction")
types3<-c("Endothelial", "Smooth_Muscle")
types4<-c("Mesenchymal_Stem_Cell","Neuronal")
types5<-c("Satellite_Cell","Macrophage")
types6<-c("T_cell", "Adipocyte")


df<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype.Rds") %>% filter(modality=="rna+atac")
df$trait<-factor(df$trait, levels=c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "INSFAST", "INS0", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE"))
print(dotplot_celltype_rnaatac(df,types1,0.05,-.5,.5, 4))
print(dotplot_celltype_rnaatac(df,types2,0.05,-.5,.7, 4))
print(dotplot_celltype_rnaatac(df,types3,0.05,-.5,.5, 4))
print(dotplot_celltype_rnaatac(df,types4,0.05,-.5,.5, 4))
print(dotplot_celltype_rnaatac(df,types5,0.05,-.5,.5, 4))
print(dotplot_celltype_rnaatac(df,types6,0.05,-.5,.5, 4))

```




```{r write_rnaatac_invnorm, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, asis=TRUE}

ds<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype.Rds") %>% 
	filter(modality=="rna+atac") %>% dplyr::select(-modality)
write.table(ds, "output/nuclei_nb/rnaatac.nuclei.invnormtrait.results.txt", col.names=T, row.names=F, quote=F, sep="\t")

```





##  Adjusting for log(fasting insulin)  
Model celltype nuclei ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + log(S_Insu) + x + offset(log(tot.nuc)   
Left figure is basic analysis; right is adjusting for log(fasting insulin)  
```{r make_nb_output_invnorm_ins, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}

out<-data.frame()
trait.ds<-readRDS("data/trait.ds.281.Rds") 
df<-readRDS("data/df.both.nuc.Rds") %>% dplyr::select(-all) 


for (j in 1:length(celltypes)) {
   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.both<-run_nb_invnorm_ins(df, traits[i], traits2[i], celltypes[j]) %>% 
		mutate(modality="rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   if (j==1) {out<-tempout} else {out<- rbind.data.frame(out, tempout) }
}
saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype_adjIns.Rds")
```


```{r dot_rna_atac_all_adjins, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each invnorm trait for all celltypes,left is basic analysis, right is same except adjusting for log(fasting insulin) in model"}

#  Plots for base
df<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype.Rds") %>% dplyr::filter(modality=="rna+atac") %>% dplyr::filter(!(trait %in% c("INSFAST", "INS0"))) 
trtlevels.ins<-c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE")
df$trait<-factor(df$trait, levels=trtlevels.ins)

types1<-c("Type_1", "Type_2a")
types2<-c("Type_2x", "Neuromuscular_junction")
types3<-c("Endothelial", "Smooth_Muscle")
types4<-c("Mesenchymal_Stem_Cell","Neuronal")
types5<-c("Satellite_Cell","Macrophage")
types6<-c("T_cell", "Adipocyte")

p1<-dotplot_celltype_rnaatac(df,types1,0.05,-.5,.5, 4)
p2<-dotplot_celltype_rnaatac(df,types2,0.05,-.5,.7,4)
p3<-dotplot_celltype_rnaatac(df,types3,0.05,-.4,.4,4)
p4<-dotplot_celltype_rnaatac(df,types4,0.05,-.5,.4,4)
p5<-dotplot_celltype_rnaatac(df,types5,0.05,-.5,.5,4)
p6<-dotplot_celltype_rnaatac(df,types6,0.05,-.5,.5,4)

#  Plots for adjins
df1<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_cell_fdrpercelltype_adjIns.Rds") %>% dplyr::filter(modality=="rna+atac") %>% dplyr::filter(!(trait %in% c("INSFAST","INS0"))) 
df1$trait<-factor(df1$trait, levels=trtlevels.ins)

p1.i<-dotplot_celltype_rnaatac(df1,types1,0.05,-.5,.5, 4)
p2.i<-dotplot_celltype_rnaatac(df1,types2,0.05,-.5,1.2, 4)
p3.i<-dotplot_celltype_rnaatac(df1,types3,0.05,-.4,.4, 4)
p4.i<-dotplot_celltype_rnaatac(df1,types4,0.05,-.6,.3, 4)
p5.i<-dotplot_celltype_rnaatac(df1,types5,0.05,-.5,1, 4)
p6.i<-dotplot_celltype_rnaatac(df1,types6,0.05,-.5,.5, 4)

print(grid.arrange(p1, p1.i, nrow=1, ncol=2))
print(grid.arrange(p2, p2.i, nrow=1, ncol=2))
print(grid.arrange(p3, p3.i, nrow=1, ncol=2))
print(grid.arrange(p4, p4.i, nrow=1, ncol=2))
print(grid.arrange(p5, p5.i, nrow=1, ncol=2))
print(grid.arrange(p6, p6.i, nrow=1, ncol=2))


```
