---
title: "Make pseudobulk counts files for hg38, also covariate files.  Note pbulk filtering to 5 counts in 25% of samples is done in 4.0_make_bulk_dataset_hg38.Rmd"
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, dev=c('png', 'pdf'), include=FALSE}
knitr::opts_chunk$set(dev = c('png', 'pdf'))
knitr::opts_chunk$set(autodep = TRUE)
```

```{r include=F}
.libPaths("/net/snowwhite/home/aujackso/R/x86_64-pc-linux-gnu-library/4.0");
library(dplyr)
library(readr)
library(stringr)
library(ggplot2)
options(scipen=1, digits=8)

```
Notes from Arushi about the RNA data  
We don't estimate contamination using genotypes per se. We filter RNA nuclei at several stages  

1. identify #UMI, mitochondrial fraction thresholds and initial qc
2. dropletUtils testEmptyDrops initially to separate nuclei from ambient
3. two-stage Demuxlet to identify singlets
4. Decontx to adjust RNA counts accounting for remaining ambient contamination. This also gives a contamination estimate and we additionally qc out nuclei here.  

 
Make snRNA counts datasets for each celltype and pseudobulk RNA.  
ATAC per person datasets are already made by Arushi.  


Note that snRNA and snATAC counts datasets were created using this code from ~/snRNAsnATAC_paper1/analysis/2.0_make_snRNAcount_ds.Rmd 
I am using links to the resulting files 
~/snRNAsnATAC_paper1/output/decontx_counts_[CELLTYPE].txt.gz
~/snRNAsnATAC_paper1/output/atac.counts.no0.[CELLTYPE].txt

These count files still contain too many samples, so must filter later.


```{r make_pbulk_counts, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
celltype<-c("Type_1", "Type_2a", "Type_2x", "Adipocyte", "Endothelial", "Macrophage", "Mesenchymal_Stem_Cell", "Neuromuscular_junction", "Neuronal", "Satellite_Cell", "Smooth_Muscle", "T_cell")

who.use.rna<-readRDS("data/df.rna.nuc.Rds")%>% pull(labelcode) 
pb<-read_tsv("data/decontx_counts_Type_1.txt.gz") %>%as.data.frame() %>% select(gene, all_of(who.use.rna))
for (i in 2:length(celltype)) {
     ds<-read_tsv(paste0("data/decontx_counts_",celltype[i],".txt.gz"))
     who.cell<-intersect(colnames(ds)[-1], who.use.rna)
     for(k in 1:length(who.cell)) {
        temp<-ds[,who.cell[k]]
	old<-pb[,who.cell[k]]
	cur<-old+temp
	pb[,who.cell[k]]<-cur
     }
}
write.table(pb, paste0("data/pbulk_counts.allgenes.txt"), col.names=T, row.names=F, quote=F, sep="\t")

```


###################################################################
#  PSEUDOBULK  
```{r prep_data_pbulk, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
pbulk<-read_tsv("data/pbulk_counts.allgenes.txt") 
	#58962
pbulk.gene<-pbulk$gene
ct.pbulk<-pbulk %>% select(-gene) %>% as.matrix()
rownames(ct.pbulk)<-pbulk.gene


#phenotypes and covariates
traits<-c("SEX", "AGE", "T2D", "INS", "BMI",  "GLU", "HEIGHT", "WEIGHT", "WAIST", "WHR", "HIP", "S_ALAT", "S_hs_CRP", "fS_Kol", "fS_Kol_HDL", "fS_Kol_LDL_c", "fS_Trigly", "GL0", "GL30", "GL60", "GL120", "S_LipoA1", "S_LipoB", "fS_C_pept", "fS_Krea", "S_GT", "S_Uraat", "fS_C_pept_30", "S_Insu_30", "B_GHb_A1C", "glu_2h_biopsy", "daily_energy_expenditure", "total_daily_physical_activity", "moderate_physical_activity", "strenuous_physical_activity",  "p_insu", "p_insu_30", "p_insu_60", "p_insu_120", "RFM", "sbp", "dbp", "ApoB_A1_ratio", "Ins_AUC_0to30", "Glu_AUC_0to30", "InsSec30", "InsGenIn", "DIo", "CpepGenIn", "HOMA", "matsuda_4pt", "matsuda_3pt", "LTPA_duration_all", "LTPA_duration_cond", "LTPA_duration_noncond", "LTPA_energy_all", "LTPA_energy_cond", "LTPA_energy_noncond", "LTPA_energy_light", "LTPA_energy_moderate", "LTPA_energy_modvig", "LTPA_energy_vigorous", "sleep_biopsy", "totalcw", "sleep_24h")      

phen0<-read_tsv("data/pbulk_covars.txt") %>% mutate(snrna_batch=paste0("B",batch)) %>% select(-batch) 

heather<-read_csv("data/tissue.csv") %>% 
	dplyr::rename(AGE=age_biopsy, INS=S_Insu, BMI=bmi, WHR=whr, GLU=glu_fast_biopsy) %>% 
	mutate(T2D=ifelse(ogtt_status_paper=="T2D",1,ifelse(ogtt_status_paper=="NGT",0,NA))) %>% 
	select(labelcode, area_name, all_of(traits))
	
p.pbulk<- merge(phen0, heather, by="labelcode")


# read in muscle tissue proportions based on snRNA + snATAC nuclei, not including muscle fiber mixed
prop<-readRDS("data/prop.both.nuc.Rds")  
p.pbulk<-merge(p.pbulk,prop,by="labelcode")


ct.pbulk<-as.data.frame(ct.pbulk) %>% dplyr::select(all_of(p.pbulk$labelcode)) %>% as.matrix()

# 58962 genes, 279 people
#check labelcode order is correct
print(all.equal(colnames(ct.pbulk), p.pbulk$labelcode))
saveRDS(p.pbulk,"data/phen_pbulk.Rds")
saveRDS(ct.pbulk, "data/ct.pbulk.Rds")  # NOT FILTERED
```


Make covariate datasets for pbulk - also RNA and ATAC

For ATAC, just use per sample per celltype tss_enrichment from /net/csgspare2/spare1/muscle_singlenuclei/clustering_f2_09_20/cluster-sample-ataqv-stats.txt  Totals are PER CELL TYPE.  

```{r make_covars_pbulk, echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE, cache=FALSE}

fusall = read_tsv("data/cluster_info_qc.tsv", guess_max=480000) %>% 
	filter(cohort=="FUSION") %>%
	dplyr::rename(labelcode=SNG.1ST) %>%
	filter(coarse_cluster_name != "Muscle_Fiber_Mixed")
	

#Pbulk 279 samples
who.use.rna<-readRDS("data/df.rna.nuc.Rds")%>% pull(labelcode) 

   ds0<-fusall %>% filter(modality=="rna") %>% 
   	filter(labelcode %in% all_of(who.use.rna)) 
   nuc0<-ds0 %>% group_by(labelcode) %>% 
	summarise(tot_nuc=n()) 
   umi0<-ds0 %>% group_by(labelcode) %>% 
	summarise(tot_umi=sum(hqaa_umi))
   mito0<-ds0 %>% group_by(labelcode) %>% 
	summarise(med.frac.mito=median(fraction_mitochondrial))
   batch0<-ds0 %>% group_by(labelcode) %>% 
	summarise(batch=unique(batch))

   covs0<- nuc0 %>% left_join(umi0) %>% left_join(mito0) %>% left_join(batch0)  %>% mutate(celltype="pbulk")

cts<-readRDS("data/ct.pbulk.Rds") 
tot.decontx.counts<-colSums(cts)
labelcode<-names(tot.decontx.counts)
cts<-cbind.data.frame(labelcode, tot.decontx.counts)	
covs.out<-covs0 %>% inner_join(cts, by="labelcode")

write.table(covs.out, "data/pbulk_covars.txt" , col.names=T, row.names=F, quote=F, sep="\t")



# RNA - 279 samples
   nuc0<-ds0 %>% group_by(labelcode) %>% 
	summarise(tot_nuc_allcelltypes=n()) 
   rna.covs0<- nuc0 %>% left_join(batch0)
write.table(rna.covs0, "data/rna_covars.txt" , col.names=T, row.names=F, quote=F, sep="\t")

  
#ATAC - 281 samples
who.use.atac<-readRDS("data/df.atac.nuc.Rds")%>% pull(labelcode) 

   ds0<-fusall %>% filter(modality=="atac") %>% 
   	filter(labelcode %in% all_of(who.use.atac)) 
   nuc0<-ds0 %>% group_by(labelcode) %>% 
	summarise(tot_nuc_allcelltypes=n()) 

ds<-read_tsv("data/atac_sample_level_covariates.tsv", guess_max=3658) %>% 
  mutate(celltype=gsub("fusion.","",coarse_cluster_name)) %>%
  filter(celltype != "Muscle_Fiber_Mixed") %>% 
  rename(labelcode=SNG.1ST) %>% 
  filter(labelcode %in% all_of(who.use.atac)) 


ds<- merge(ds, nuc0, by="labelcode")
write.table(ds, "data/atac_covars.txt", col.names=T, row.names=F, quote=F, sep="\t")

table(ds$celltype)
# 3350 rows = 8 celltypes with 281 samples, Adipocyte with 280, NMJ with 278, Satellite with 276, Macrophage with 268

```

```{r prep_data_pbulk, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}
pbulk<-read_tsv("data/pbulk_counts.allgenes.txt") 
	#58962
pbulk.gene<-pbulk$gene
ct.pbulk<-pbulk %>% select(-gene) %>% as.matrix()
rownames(ct.pbulk)<-pbulk.gene


#phenotypes and covariates
traits<-c("SEX", "AGE", "T2D", "INS", "BMI",  "GLU", "HEIGHT", "WEIGHT", "WAIST", "WHR", "HIP", "S_ALAT", "S_hs_CRP", "fS_Kol", "fS_Kol_HDL", "fS_Kol_LDL_c", "fS_Trigly", "GL0", "GL30", "GL60", "GL120", "S_LipoA1", "S_LipoB", "fS_C_pept", "fS_Krea", "S_GT", "S_Uraat", "fS_C_pept_30", "S_Insu_30", "B_GHb_A1C", "glu_2h_biopsy", "daily_energy_expenditure", "total_daily_physical_activity", "moderate_physical_activity", "strenuous_physical_activity",  "p_insu", "p_insu_30", "p_insu_60", "p_insu_120", "RFM", "sbp", "dbp", "ApoB_A1_ratio", "Ins_AUC_0to30", "Glu_AUC_0to30", "InsSec30", "InsGenIn", "DIo", "CpepGenIn", "HOMA", "matsuda_4pt", "matsuda_3pt", "LTPA_duration_all", "LTPA_duration_cond", "LTPA_duration_noncond", "LTPA_energy_all", "LTPA_energy_cond", "LTPA_energy_noncond", "LTPA_energy_light", "LTPA_energy_moderate", "LTPA_energy_modvig", "LTPA_energy_vigorous", "sleep_biopsy", "totalcw", "sleep_24h")      

phen0<-read_tsv("data/pbulk_covars.txt") %>% mutate(snrna_batch=paste0("B",batch)) %>% select(-batch) 

heather<-read_csv("data/tissue.csv") %>% 
	dplyr::rename(AGE=age_biopsy, INS=S_Insu, BMI=bmi, WHR=whr, GLU=glu_fast_biopsy) %>% 
	mutate(T2D=ifelse(ogtt_status_paper=="T2D",1,ifelse(ogtt_status_paper=="NGT",0,NA))) %>% 
	select(labelcode, area_name, all_of(traits), ogtt_status_paper)
	
p.pbulk<- merge(phen0, heather, by="labelcode")


# read in muscle tissue proportions based on snRNA + snATAC nuclei, not including muscle fiber mixed
prop<-readRDS("data/prop.both.nuc.Rds")  
p.pbulk<-merge(p.pbulk,prop,by="labelcode")


ct.pbulk<-as.data.frame(ct.pbulk) %>% dplyr::select(all_of(p.pbulk$labelcode)) %>% as.matrix()

# 58962 genes, 279 people
#check labelcode order is correct
print(all.equal(colnames(ct.pbulk), p.pbulk$labelcode))
saveRDS(p.pbulk,"data/phen_pbulk.Rds")
saveRDS(ct.pbulk, "data/ct.pbulk.Rds")  # NOT FILTERED
```




