---
title: "Descriptive statistics and basic QC for nuclei. Make proportions for downstream analysis."
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, dev=c('png'), include=F}
knitr::opts_chunk$set(dev = c('png'))
knitr::opts_chunk$set(autodep = TRUE)

```


```{r include=F}
.libPaths("/net/snowwhite/home/aujackso/R/x86_64-pc-linux-gnu-library/4.0");
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(scales)
library(kableExtra)
library(skimr)
library(RColorBrewer)
options(scipen=1, digits=5)
options(nwarnings=10000)

invnorm = function(x, seed) {
  set.seed(seed)
  qnorm((rank(x, na.last="keep", ties.method="random") - 0.5) / sum(!is.na(x)))
}
vio<-function(ds, x,y, xname, yname, titl, palet) {
	return(
		ggplot(ds, aes(x=x, y=y))  + 
		geom_jitter(aes(color=x), shape=16, alpha=0.7, position=position_jitter(width=0.2,height=0), na.rm=T) + 
		geom_violin(aes(color=x), size=1, alpha=.5, na.rm=T)  +  
		geom_boxplot(notch=T, width=0.5, alpha=.5, na.rm=T)  +  
		scale_colour_manual(values=palet)  + 
		theme_bw() + 
		theme(legend.position="none") + 
		theme(axis.text.x=element_text(angle=0, size=10)) + 
		theme(axis.text.y=element_text(size=10)) + 
		xlab(paste0(xname)) + ylab(paste0(yname)) +
		ggtitle(label=titl) + 
		theme(plot.title = element_text(size=14)) + 
		facet_wrap(~celltype, scales="free") +
	theme(strip.text.x=element_text(size=8)) + 
		xlab(paste0(xname))
	)	
}
vio.log<-function(ds, x,y, xname, yname, titl,palet) {
	return(
		ggplot(ds, aes(x=x, y=y))  + 
		geom_jitter(aes(color=x), shape=16, alpha=0.7, position=position_jitter(width=0.2,height=0), na.rm=T) + 
		geom_violin(aes(color=x), size=1, alpha=.5, na.rm=T)  +  
		geom_boxplot(notch=T, width=0.5, alpha=.5, na.rm=T)  +  
		scale_colour_manual(values=palet)  + 
		theme_bw() + 
		scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x),labels = function(x)round(x,-1)) + 
		theme(legend.position="none") + 
		theme(axis.text.x=element_text(angle=0, size=10)) + 
		theme(axis.text.y=element_text(size=10)) + 
		xlab(paste0(xname)) + ylab(paste0(yname)) +
		ggtitle(label=titl) + 
		theme(plot.title = element_text(size=14)) + 
		facet_wrap(~celltype, scales="free") +
	theme(strip.text.x=element_text(size=8)) + 
		xlab(paste0(xname))
	)	
}
vio_sing.log<-function(ds, x,y, xname, yname, titl,palet) {
	return(
		ggplot(ds, aes(x=x, y=y))  + 
		geom_jitter(aes(color=x), shape=16, alpha=0.7, position=position_jitter(width=0.2,height=0), na.rm=T) + 
		geom_violin(aes(color=x), size=1, alpha=.5, na.rm=T)  +  
		geom_boxplot(notch=T, width=0.5, alpha=.5, na.rm=T)  +  
		scale_colour_manual(values=palet)  + 
		theme_bw() + 
		scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x),labels = function(x)round(x,-1)) + 
		theme(legend.position="none") + 
		theme(axis.text.x=element_text(angle=0, size=10)) + 
		theme(axis.text.y=element_text(size=10)) + 
		xlab(paste0(xname)) + ylab(paste0(yname)) +
		ggtitle(label=titl) + 
		theme(plot.title = element_text(size=14)) + 
		xlab(paste0(xname))
	)	
}



# colors from Arushi's cluster_colors.tsv file in /net/csgspare2/spare1/muscle_singlenuclei/hg38/fusion_plus_multiome_cluster_info/
Mycolors = c("#a6cee3", "#1f78b4", "#8dd3c7", "#b2df8a",
           "#6a3d9a","#fb9a99", "#fdbf6f", "#cab2d6", "#b15928", "#e31a1c", 
           "#ffff33","#999999")
cell.levels<-c("Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction",  "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")

#  NOTE  Fibro-adipogenic progenitors is the new term for Mesenchymal_Stem_Cell

# Paired color palette + teal + grey for all, for 14 colors total
Mycolors1 = c("#808080", "#a6cee3", "#1f78b4", "#8dd3c7", "#b2df8a", 
           "#fb9a99", "#fdbf6f", "#cab2d6", "#b15928", "#e31a1c", 
           "#ffff33","#999999")
cell.levels1<-c("all", "Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction",  "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")


cbPalette <- c("#CCCCCC", "#CFD7F8", "#F0E442","#56B4E9", "#009E73", "#D55E00", "#CC79A7", "#0072B2")
#               grey,       lilac,   yellow,    ltblue,    green,     red,       pink,       blue 

# note proportions will include NAs if a sample did not have any nuclei for the celltype
zscore = function(x) { (x-mean(x, na.rm=T))/sd(x, na.rm=T); }
make_props<-function(ds) {
	samptot<-ds %>% select(-labelcode) %>% rowSums(.,na.rm=T)
	temp<-ds %>% select(-labelcode) %>% as.matrix()
	prop=temp/samptot
	labelcode<-ds$labelcode
	out.prop<-cbind.data.frame(labelcode, prop)
	prop.z = data.frame(apply(prop, 2, zscore), check.names=F)  
	out.prop.z<-cbind.data.frame(labelcode, prop.z)
	out<-list()
	out[[1]]<-out.prop
	out[[2]]<-out.prop.z
	return(out)
}

boxp<-function(ds) {
return(ggplot(ds, aes(x=trait, y=value, color=group))  + 
      geom_boxplot(size=0.1, alpha=.5, color="black")  +  
        geom_point(alpha=0.3, size=4)  + 
	geom_text(aes(label=txt, hjust=0.1, angle=-30)) + 
        scale_colour_manual(values=cbPalette)  + 
        theme_bw() + 
        theme(axis.text.x=element_text(angle=45, size=8, hjust=1)) + 
        theme(axis.text.y=element_text(size=15)) 
)
}


jitboxp<-function(ds,trtstr, palet) {
return(ggplot(ds, aes(x=trait, y=value, color=group))  + 
      geom_boxplot(size=0.1, alpha=.5, color="black")  +  
        #geom_point(alpha=0.3, size=4)  + 
	geom_jitter(alpha=1,size=4, position=position_jitter(width=0.2,height=0)) + 
	geom_text(aes(label=txt, hjust=-0.05, angle=-30)) + 
        scale_colour_manual(values=palet)  + 
        theme_bw() + 
        theme(axis.text.x=element_blank()) + 
        theme(axis.text.y=element_text(size=15)) + 
	xlab(paste0(trtstr)) + ylab(paste0(trtstr))  
 
)
}

histos<-function(ds, xstr) {
	tp<- gather(ds, key="celltype", value="val",-labelcode)
	tp$celltype<-factor(tp$celltype, levels=c("Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction", "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")) 
	return(ggplot(tp, aes(val, color=celltype)) +  
		geom_histogram(show.legend=FALSE,fill="gray90",na.rm=T) +
		scale_colour_manual(values=Mycolors) +
		theme(axis.text.x=element_text(angle=45, size=10)) + 
		facet_wrap(~celltype, scales="free") +
		xlab(paste0(xstr))
	)	
}		 

histos_logx<-function(ds, xstr) {
	tp<- gather(ds, key="celltype", value="val",-labelcode)
	tp$val<-tp$val+1
	tp$celltype<-factor(tp$celltype, levels=c("Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction", "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")) 
	return(ggplot(tp, aes(val, color=celltype)) +  
		geom_histogram(show.legend=FALSE,fill="gray90",na.rm=T) +
		scale_colour_manual(values=Mycolors) +
		scale_x_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x),labels = function(x)round(x,-1)) + 
		theme(axis.text.x=element_text(angle=45, size=10)) + 
		facet_wrap(~celltype, scales="free") +
		xlab(paste0(xstr))
	)	
}		 

```

* Drop nuclei for samples
** [1 sample]:  non-Finnish (METSIM MHUC-901, born in France, 'finnish'=0)
** [1 sample]:  female NGT 1st degree relative of retained female IGT 22012
** [1 sample]:  female NGT 1st degree relative of retained female IGT 32074
* For overall nuclei counts and proportions - DROP SAMPLES WITH <100 NUCLEI IN EITHER RNAseq or ATACseq ie 2 samples with <100 RNA nuclei - 282 remain   
* For RNA nuclei counts - Drop 2 samples with <100 RNA nuclei - 282 remain 
* For ATAC nuclei counts - Drop 0 samples - 284 remain   
 
```{r level1_qc, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE, results='asis'}
# cluster_info_qc.tsv format
#index	UMAP_1	UMAP_2	modality	library	barcode	batch	cluster	demux_cat	SNG.1ST	hqaa_umi	fraction_mitochondrial	cohort	sample	tss_enrichment	log10_hqaa_umi	coarse_cluster_name
#rna.3172.NM.1.AAACAGCCAAGCCACT	2.42076191597117	5.496282517399349	rna	3172-NM-1	AAACAGCCAAGCCACT	3172	0	multiome	multiome	2974.0	0.008721516434765073	multiome	multiome--3172	NA	7.9976631270201	Type_1

# modality: 188337 "rna" and 268543 "atac" nuclei including 15508 multiome nuclei.  

# 14 clusters but 13 coarse cluster names: cluster 0 is Type1, clusters 3 and 5 are Endothelial.  

# demux_cat =   7754 multiome,   7754 NA,  12499 nS-SNG, 260789 SNG,  61679 SNG-nS, 106405 SNG-SNG

# cohort: 441372 FUSION, 15508 multiome
# 

ds.orig = read_tsv("/net/csgspare2/spare1/muscle_singlenuclei/hg38/fusion_plus_multiome_cluster_info/cluster_info_qc.tsv", guess_max=480000) %>% 
   	rename(labelcode=SNG.1ST) %>% 
	filter(cohort=="FUSION") 
length(unique(ds.orig$labelcode)) #287 FUSION samples originally	
ds0 <- ds.orig %>%
	filter(!(labelcode %in% c(omittedsampleids))) %>% 
	filter(coarse_cluster_name != "Muscle_Fiber_Mixed")

length(unique(ds0$labelcode)) #284 FUSION samples	



all0<-ds0 %>% 	 
	group_by(labelcode) %>% summarise(nuc=n())
colnames(all0)<-c("labelcode", "all.nuc")
rna0<-ds0 %>% filter(modality=="rna") %>%
	group_by(labelcode) %>% summarise(nuc=n())
colnames(rna0)<-c("labelcode", "rna.nuc")
atac0<-ds0 %>% filter(modality=="atac") %>%
	group_by(labelcode) %>% summarise(nuc=n())
colnames(atac0)<-c("labelcode", "atac.nuc")
all0<- all0 %>% left_join(rna0, by="labelcode") %>%
	left_join(atac0, by="labelcode") %>% arrange(all.nuc)

#  Dropping samples that don't have at least 100 nuclei in either RNA or ATAC
any.ds<- all0 %>% filter(rna.nuc>=100 | atac.nuc>=100)
nrow(any.ds) # 281 samples as our baseline dataset

#  USING at least 100 RNA and at least 100 ATAC nuclei as the paper dataset
both.ds  <- all0 %>% filter(rna.nuc>=100 & atac.nuc>=100) 
nrow(both.ds) #279
rna.ds <- all0 %>% filter(rna.nuc>=100)  
nrow(rna.ds) #279
atac.ds <- all0 %>% filter(atac.nuc>=100)  
nrow(atac.ds) #281



#####################################################################
#  Check to see if numbers of samples change if we had not taken out muscle fiber mixed at the start...
ds1 <- ds.orig %>%
	filter(!(labelcode %in% c("12004", "22011", "32071"))) 
all1<-ds1 %>% 	 
	group_by(labelcode) %>% summarise(nuc=n())
colnames(all1)<-c("labelcode", "all.nuc")
rna1<-ds1 %>% filter(modality=="rna") %>%
	group_by(labelcode) %>% summarise(nuc=n())
colnames(rna1)<-c("labelcode", "rna.nuc")
atac1<-ds1 %>% filter(modality=="atac") %>%
	group_by(labelcode) %>% summarise(nuc=n())
colnames(atac1)<-c("labelcode", "atac.nuc")
all1<- all1 %>% left_join(rna1, by="labelcode") %>%
	left_join(atac1, by="labelcode") %>% arrange(all.nuc)

#  Dropping samples that don't have at least 100 nuclei in either RNA or ATAC
any.ds1<- all1 %>% filter(rna.nuc>=100 | atac.nuc>=100)
nrow(any.ds1) # 281 samples as our basic dataset

both.ds1  <- all1 %>% filter(rna.nuc>=100 & atac.nuc>=100) 
nrow(both.ds1) #279
rna.ds1 <- all1 %>% filter(rna.nuc>=100)  
nrow(rna.ds1) #279
atac.ds1 <- all1 %>% filter(atac.nuc>=100)  
nrow(atac.ds1) #281

#  numbers didn't change if added MFM nuclei

```


```{r make_df_nuc1, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide'}

celltype<-c("Type_1", "Type_2a", "Type_2x", "Adipocyte", "Endothelial", "Macrophage", "Mesenchymal_Stem_Cell",  "Neuromuscular_junction", "Neuronal", "Satellite_Cell", "Smooth_Muscle", "T_cell")


# Start with 281 samples
print(dim(ds0))

# nuclei count dataset
make_df_nuc<-function(df0, type) {

	if (type=="rna") { ds <- df0 %>% filter(modality=="rna") }
	if (type=="atac") { ds <- df0 %>% filter(modality=="atac") }
	if (type=="any") { ds <- df0  }
	out<-data.frame()
	temp<-ds %>% 
		group_by(labelcode) %>% summarise(nuc=n())
	colnames(temp)<-c("labelcode", "all")
	out<-temp
	for (i in 1:length(celltype)) {
		temp<-ds %>% filter(coarse_cluster_name==celltype[i]) %>% 
			group_by(labelcode) %>% summarise(nuc=n())
		colnames(temp)<-c("labelcode", celltype[i])
		out<-merge(out,temp,by="labelcode",all=T)
	}	
	out[is.na(out)] <- 0
	return(out)
}

df.rna.nuc<-make_df_nuc(ds0, "rna") %>% filter(all>=100) #279
df.atac.nuc<-make_df_nuc(ds0, "atac") %>% filter(all>=100) #281
df.any.nuc<-make_df_nuc(ds0, "any")  #281
df.both.nuc<-make_df_nuc(ds0, "any") %>% filter(labelcode %in% all_of(df.rna.nuc$labelcode)) #279
df.any.nuc<-make_df_nuc(ds0, "any")  #281

saveRDS(df.rna.nuc, "data/df.rna.nuc.Rds") #279 samples
saveRDS(df.atac.nuc, "data/df.atac.nuc.Rds") #281 samples
saveRDS(df.both.nuc, "data/df.both.nuc.Rds") #279 samples
saveRDS(df.any.nuc, "data/df.any.nuc.Rds") #281 samples

```


RNA + ATAC nuclei counts for 281 samples    
```{r violin_nuclei_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='asis', fig.cap="counts of RNA+ATAC nuclei, per cell type"}
ds<- df.any.nuc 
tp0<- gather(ds, key="celltype", value="any.nuc",-labelcode)
tp0$celltype<-factor(tp0$celltype, levels=cell.levels1) 

temp.all<-tp0 %>% filter(celltype=="all")
print(vio_sing.log(temp.all, temp.all$celltype, temp.all$any.nuc, "overall", "rna+atac #nuclei", "Overall # nuclei per 281 samples, log scale", Mycolors1))



cat("List of samples with <400 RNA+ATAC nuclei  \n")
temp<-ds %>% filter(all<400) %>% arrange(all) %>% select(labelcode, all)
temp.who <- temp$labelcode
temp.rna<- df.rna.nuc %>% rename(rna=all) %>% select(labelcode, rna) %>% filter(labelcode %in% all_of(temp.who))
temp.atac<- df.atac.nuc %>% rename(atac=all) %>% select(labelcode, atac) %>% filter(labelcode %in% all_of(temp.who))
temp<-merge(temp, temp.rna, by="labelcode") 
temp<-merge(temp, temp.atac, by="labelcode") %>% arrange(all)
kable(temp) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F)

```

Look at proportion outliers. 
```{r level3_qc, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='asis',fig.cap="nuclei counts, per cell type" }
# Proportions based on any ATAC + RNA nuclei - just to check Z's of samples not included in RNA
df.any<-df.any.nuc %>% select(-all)	
list.any.nuc<-make_props(df.any)
prop.any.nuc<-list.any.nuc[[1]]
prop.any.nuc.z<-list.any.nuc[[2]] 
summary(prop.any.nuc.z)
temp<-prop.any.nuc.z %>% filter(Adipocyte>5 | Endothelial > 5 | Macrophage > 5 | Mesenchymal_Stem_Cell>5 | Neuromuscular_junction>5 | Neuronal>5 | Smooth_Muscle>5 | T_cell>5)
temp


# Proportions based on ATAC nuclei
df.atac<-df.atac.nuc %>% select(-all)	
list.atac.nuc<-make_props(df.atac)
prop.atac.nuc<-list.atac.nuc[[1]]
prop.atac.nuc.z<-list.atac.nuc[[2]] 
summary(prop.atac.nuc.z)
temp<-prop.atac.nuc.z %>% filter(Adipocyte>5 | Endothelial > 5 | Macrophage > 5 | Mesenchymal_Stem_Cell>5 | Neuromuscular_junction>5 | Neuronal>5 | Smooth_Muscle>5 | T_cell>5)
temp

# repeat for RNA nuclei only
df.rna<-df.rna.nuc %>% select(-all)	
list.rna.nuc<-make_props(df.rna)
prop.rna.nuc<-list.rna.nuc[[1]]
prop.rna.nuc.z<-list.rna.nuc[[2]] 
summary(prop.rna.nuc.z)
temp<-prop.rna.nuc.z %>% filter(Adipocyte>5 | Endothelial > 5 | Macrophage > 5 | Mesenchymal_Stem_Cell>5 | Neuromuscular_junction>5 | Neuronal>5 | Smooth_Muscle>5 | T_cell>5)
temp
 
df.both<-df.both.nuc %>% select(-all)	
list.both.nuc<-make_props(df.both)
prop.both.nuc<-list.both.nuc[[1]]
prop.both.nuc.z<-list.both.nuc[[2]] 
summary(prop.both.nuc.z)
temp<-prop.both.nuc.z %>% filter(Adipocyte>5 | Endothelial > 5 | Macrophage > 5 | Mesenchymal_Stem_Cell>5 | Neuromuscular_junction>5 | Neuronal>5 | Smooth_Muscle>5 | T_cell>5)
temp
 
saveRDS(prop.both.nuc, "data/prop.both.nuc.Rds") #includes lots of 0s


```

Box plots of nuclei counts (log scale)  
Labelling is z-scores of the LOG(total nuclei counts per modality).  

```{r qc_box, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='asis',fig.cap="nuclei counts, per cell type" }

#Box plots of nuclei counts per modality
temp.both<-df.both.nuc %>% mutate(group="both") %>% mutate(z=round(zscore(log(all)), digits=2)) %>% select(labelcode, group, all, z)
temp.rna<-df.rna.nuc %>% mutate(group="rna") %>% mutate(z=round(zscore(log(all)), digits=2)) %>% select(labelcode, group, all, z)
temp.atac<-df.atac.nuc %>% mutate(group="atac") %>% mutate(z=round(zscore(log(all)), digits=2)) %>% select(labelcode, group, all, z)
temp<-rbind.data.frame(temp.both, temp.rna, temp.atac) 
temp$group<-factor(temp$group, levels=c("both", "atac", "rna"))
print(ggplot(temp, aes(x=group, y=all, color=group))  + 
      geom_boxplot(size=0.2, alpha=.7, color="black")  +  
        #geom_point(alpha=0.3, size=4)  + 
	geom_jitter(alpha=0.5,size=2, position=position_jitter(width=0.2,height=0)) + 
	geom_text(aes(label=z, hjust=-1.0, angle=-30)) + 
        #scale_colour_manual(values=cbPalette)  + 
	scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x),labels = function(x)round(x,-1)) + 
        theme_bw() + 
        theme(axis.text.x=element_blank()) + 
        theme(axis.text.y=element_text(size=15)) + 
	xlab("modality") + ylab("number of nuclei (log scale)")  
) 

```

## How many samples are outliers for cell type proportions (|z>5|) made from RNA+ATAC nuclei?  
```{r violin_nuclei_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='asis', fig.cap="RNA+ATAC nuclei counts, per cell type"}

tp0<- gather(df.both.nuc, key="celltype", value="rna.atac.nuc",-labelcode, -all)
tp0$celltype<-factor(tp0$celltype, levels=cell.levels) 
tp0$nuc<-tp0$rna.atac.nuc+1
print(vio.log(tp0, tp0$celltype, tp0$nuc, "celltype", "RNA+ATAC nuclei", "RNA+ATAC nuclei counts +1 per sample, log scale", Mycolors))
tp0<- tp0 %>% select(-nuc)

tp00<- gather(as.data.frame(prop.both.nuc), key="celltype", value="prop", -labelcode)
tp00$celltype<-factor(tp00$celltype, levels=cell.levels) 
print(vio(tp00, tp00$celltype, tp00$prop, "celltype", "cell type percentage", "RNA + ATAC nuclei counts:  cell type percentages",Mycolors))


tp<- gather(as.data.frame(prop.both.nuc.z), key="celltype", value="prop.z", -labelcode)
tp$celltype<-factor(tp$celltype, levels=cell.levels) 
print(vio(tp, tp$celltype, tp$prop.z, "celltype", "Z score of cell type percentage", "RNA + ATAC nuclei counts: cell type percentage Z scores",Mycolors))

z5<-tp %>% filter(abs(prop.z)>5)
colnames(z5)<-c("labelcode", "celltype", "Z.celltype.pct")
z.rnaatac<-z5 %>% merge(tp00, by=c("labelcode", "celltype")) %>% merge(tp0, by=c("labelcode", "celltype"))
cat(paste0(length(unique(z.rnaatac$labelcode))," outliers (|z|>5) for proportions based on RNA+ATAC nuclei counts  \n"))
 kable(z.rnaatac) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F)

 z5.summary <- z5 %>% group_by(labelcode) %>% summarize(z5.both=n())
 
```
9 sample outliers for combined RNA+ATAC nuclei.  

## How many samples are outliers for z of cell type proportions made from RNA nuclei?
```{r violin_rna_nuclei_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='asis', fig.cap="RNA nuclei counts, per cell type"}

df<-df.rna.nuc %>% select(-all) 
list.rna.nuc<-make_props(df)
prop.rna.nuc<-list.rna.nuc[[1]]
prop.rna.nuc.z<-list.rna.nuc[[2]] 

tp0<- gather(df, key="celltype", value="rna.nuc",-labelcode)
tp0$celltype<-factor(tp0$celltype, levels=cell.levels) 
tp0$nuc<-tp0$rna.nuc+1
print(vio.log(tp0, tp0$celltype, tp0$nuc, "celltype", "RNA nuclei", "RNA nuclei counts +1 per sample, log scale", Mycolors))
tp0<- tp0 %>% select(-nuc)


tp00<- gather(as.data.frame(prop.rna.nuc), key="celltype", value="prop", -labelcode)
tp00$celltype<-factor(tp00$celltype, levels=cell.levels) 
print(vio(tp00, tp00$celltype, tp00$prop, "celltype", "cell type percentage", "RNA nuclei counts:  cell type percentages",Mycolors))


tp<- gather(as.data.frame(prop.rna.nuc.z), key="celltype", value="prop.z", -labelcode)
tp$celltype<-factor(tp$celltype, levels=cell.levels) 
print(vio(tp, tp$celltype, tp$prop.z, "celltype", "Z score of cell type percentage", "RNA nuclei counts: cell type percentage Z scores",Mycolors))

z5.rna<-tp %>% filter(abs(prop.z)>5)
colnames(z5.rna)<-c("labelcode", "celltype", "Z.celltype.pct")
z.rna<-z5.rna %>% merge(tp00, by=c("labelcode", "celltype")) %>% merge(tp0, by=c("labelcode", "celltype"))
#cat(paste0(length(unique(z.rna$labelcode))," outliers (|z|>5) for proportions based on RNA nuclei counts  \n")) 
#kable(z.rna) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F)
#write.table(z.rna, "output/z5.rna.outlier.celltype.txt", row.names=F, col.names=T, quote=F, sep="\t") 
 
 z5.summary.rna <- z5.rna %>% group_by(labelcode) %>% summarize(z5.rna=n())
#write.table(z5.summary.rna, "output/z5.rna.outliers.txt", row.names=F, col.names=T, quote=F, sep="\t") 

 
```
```{r violin_atac_nuclei_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='asis', fig.cap="ATAC nuclei counts, per cell type"}
df<-df.atac.nuc %>% select(-all)
list.atac.nuc<-make_props(df)
prop.atac.nuc<-list.atac.nuc[[1]]
prop.atac.nuc.z<-list.atac.nuc[[2]] 

tp0<- gather(df, key="celltype", value="atac.nuc",-labelcode)
tp0$celltype<-factor(tp0$celltype, levels=cell.levels) 
tp0$nuc<-tp0$atac.nuc+1
print(vio.log(tp0, tp0$celltype, tp0$nuc, "celltype", "ATAC nuclei", "ATAC nuclei counts +1 per sample, log scale", Mycolors))
tp0<- tp0 %>% select(-nuc)


tp00<- gather(as.data.frame(prop.atac.nuc), key="celltype", value="prop", -labelcode)
tp00$celltype<-factor(tp00$celltype, levels=cell.levels) 
print(vio(tp00, tp00$celltype, tp00$prop, "celltype", "cell type percentage", "ATAC nuclei counts:  cell type percentages",Mycolors))


tp<- gather(as.data.frame(prop.atac.nuc.z), key="celltype", value="prop.z", -labelcode)
tp$celltype<-factor(tp$celltype, levels=cell.levels) 
print(vio(tp, tp$celltype, tp$prop.z, "celltype", "Z score of cell type percentage", "ATAC nuclei counts: cell type percentage Z scores",Mycolors))

z5.atac<-tp %>% filter(abs(prop.z)>5)
colnames(z5.atac)<-c("labelcode", "celltype", "Z.celltype.pct")
z.atac<-z5.atac %>% merge(tp00, by=c("labelcode", "celltype")) %>% merge(tp0, by=c("labelcode", "celltype"))
#cat(paste0(length(unique(z.atac$labelcode))," outliers (|z|>5) for proportions based on ATAC nuclei counts  \n"))
# kable(z.atac) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F)
#write.table(z.atac, "output/z5.atac.outlier.celltype.txt", row.names=F, col.names=T, quote=F, sep="\t") 


 
```

## How many samples if we draw threshold at #nuclei per sample, per celltype?
This is for downstream phenotype analysis 
```{r celltype_threshold_nuclei, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="available samples with N nuclei per cell type"}

Mycolors = c("#a6cee3", "#1f78b4", "#8dd3c7", "#b2df8a", 
           "#6a3d9a","#fb9a99", "#fdbf6f", "#cab2d6", "#b15928", "#e31a1c", 
           "#ffff33","#999999")
celltype<-c("Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction", "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")


count_nuclei<-function(ds) {
	out<-data.frame()
	th<-c(0,1,2,3,4,5,6,7,8,9,10)
	for (i in 1:length(th)) {
	for (j in 1:length(celltype)) {
		thresh<-th[i]
		x<-ds %>% pull(celltype[j])
		n<-length(x[x>thresh & !is.na(x)])
		cell<-celltype[j]
		temp<-cbind.data.frame(cell, n, thresh)
		if(i==1 & j==1) {out<-temp}
		if(i==1 & j>1 | i>1) {out<-rbind.data.frame(out, temp)}
	}	
	}
	return(out)
}
rna.thresh<-count_nuclei(df.rna.nuc) %>% mutate(celltype=factor(cell, levels=celltype))			
atac.thresh<-count_nuclei(df.atac.nuc)	 %>% mutate(celltype=factor(cell, levels=celltype))			

plotn<-function(ds, titl) {
	return(
		ggplot(ds, aes(x=thresh, y=n, color=celltype))  + 
		geom_point() + geom_line() + 
		scale_colour_manual(values=Mycolors)  + 
		theme_bw() + 
		#theme(legend.position="none") + 
		theme(axis.text.x=element_text(angle=0, size=10)) + 
		theme(axis.text.y=element_text(size=10)) + 
		xlab(paste0("nuclei threshold")) + ylab("# samples") +
		ggtitle(label=titl) + 
		theme(plot.title = element_text(size=14)) 
	)	
}

plotn(rna.thresh, "Samples with RNA nuclei")
plotn(atac.thresh, "Samples with ATAC nuclei")


```



```{r scatter_functions, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, asis=TRUE}

#  make scatterplots of rna nuclei vs atac nuclei per sample, per celltype
scatter.all<-function(ds.rna, ds.atac, xstr, ystr ) {
	tp.rna<- gather(ds.rna, key="celltype", value="rna",-labelcode)
	tp.atac<- gather(ds.atac, key="celltype", value="atac",-labelcode)
	tp<-merge(tp.rna, tp.atac, by=c("labelcode", "celltype")) 
tp$celltype<-factor(tp$celltype, levels=c("all","Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction", "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")) 
# Paired color palette + teal + grey for all, for 14 colors total
Mycolors1 = c("#808080", "#a6cee3", "#1f78b4", "#66C2A5", "#b2df8a", 
           "#6a3d9a","#fb9a99", "#fdbf6f", "#cab2d6", "#b15928", "#e31a1c", 
           "#FFD92F","#B3B3B3")

	return(ggplot(tp, aes(x=rna, y=atac, color=celltype)) +  
		geom_point(alpha=0.4,size=2, show.legend=FALSE,na.rm=T) +
		scale_colour_manual(values=Mycolors1) + 
		geom_abline(slope=1, intercept=0, colour="black") + 
		xlab(xstr) + ylab(ystr) +  
		facet_wrap(~celltype, scales="free")  
	)	
}		 

scatter<-function(ds.rna, ds.atac, xstr, ystr ) {
	tp.rna<- gather(ds.rna, key="celltype", value="rna",-labelcode)
	tp.atac<- gather(ds.atac, key="celltype", value="atac",-labelcode)
	tp<-merge(tp.rna, tp.atac, by=c("labelcode", "celltype")) 
tp$celltype<-factor(tp$celltype, levels=c("Type_1", "Type_2a",  "Type_2x", "Neuromuscular_junction",  "Mesenchymal_Stem_Cell",  "Endothelial", "Smooth_Muscle",  "Satellite_Cell", "Macrophage", "T_cell", "Neuronal", "Adipocyte")) 
Mycolors = c("#a6cee3", "#1f78b4", "#66C2A5", "#b2df8a", 
           "#6a3d9a","#fb9a99", "#fdbf6f", "#cab2d6", "#b15928", "#e31a1c", 
           "#FFD92F","#B3B3B3")

	return(ggplot(tp, aes(x=rna, y=atac, color=celltype)) +  
		geom_point(alpha=0.4,size=2, show.legend=FALSE,na.rm=T) +
		scale_colour_manual(values=Mycolors) + 
		geom_abline(slope=1, intercept=0, colour="black") + 
		xlab(xstr) + ylab(ystr) +  
		facet_wrap(~celltype, scales="free")  
	)	
}		 


```


## Scatterplots #nuclei or cell type proportions for RNA vs ATAC  
NO OUTLIERS REMOVED BASED ON CELL TYPE PROPORTIONS  
Restricted to `r nrow(df.both.nuc2)` samples with RNA and ATAC nuclei and >100 RNA nuclei  

```{r scat_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="scatterplot of proportions of RNA vs ATAC nuclei or proportions, per cell type"}

print(scatter.all(df.rna.nuc, df.atac.nuc, "RNA #nuclei", "ATAC #nuclei"))
print(scatter(prop.rna.nuc, prop.atac.nuc, "proportion based on RNA nuclei", "proportion based on ATAC nuclei"))
print(scatter(prop.both.nuc, prop.atac.nuc, "proportion based on RNA+ATAC nuclei", "proportion based on ATAC nuclei"))
print(scatter(prop.both.nuc, prop.rna.nuc, "proportion based on RNA+ATAC nuclei", "proportion based on RNA nuclei"))

```



## Histograms of nuclei or count proportions and #nuclei for RNA, ATAC and RNA+ATAC  
Restricted to `r nrow(prop.both.nuc)` samples with RNA and ATAC nuclei and >100 RNA nuclei    

```{r histo_proportion_nuclei_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="proportions of nuclei, per cell type"}

print(histos(prop.rna.nuc, "rna proportion of nuclei"))
print(histos(prop.atac.nuc, "atac proportion of nuclei"))
print(histos(prop.both.nuc, "rna+atac proportion of nuclei"))
```

```{r histo_nuclei_per_celltype, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nuclei counts,  per cell type"}
df.rna.nuc<- df.rna.nuc2 %>% select(-all)		 
df.atac.nuc<- df.atac.nuc2 %>% select(-all)		 
df.both.nuc<- df.both.nuc2 %>% select(-all)		 
print(histos(df.rna.nuc, "rna counts of nuclei"))
print(histos(df.atac.nuc, "atac counts of nuclei"))
print(histos(df.both.nuc, "rna+atac counts of nuclei"))
```

