---
title: "Negative binomial regression for total (not cell specific) RNA + ATAC nuclei count data, using invnorm traits"
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, dev=c('png'), include=F}
knitr::opts_chunk$set(dev = c('png'))
knitr::opts_chunk$set(autodep = TRUE)

```


```{r include=F}
.libPaths("/net/snowwhite/home/aujackso/R/x86_64-pc-linux-gnu-library/4.0");
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(broom)
library(MASS)
options(scipen=1, digits=5)
options(nwarnings=10000)
library(scales)
library(grid)
library(gridExtra)
library(kableExtra)

invnorm = function(x, seed) {
  set.seed(seed)
  qnorm((rank(x, na.last="keep", ties.method="random") - 0.5) / sum(!is.na(x)))
}

```


```{r functions, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}

run_nb_invnorm<-function(ds0,trt,trt2, type) {
	print(paste0(trt, " " , type ))
	temp<- ds0 %>% dplyr::select(-labelcode) 
	tot.nuc<-rowSums(temp, na.rm=T)
	labelcode<-ds0$labelcode
	temp.totnuc<-cbind.data.frame(labelcode, tot.nuc)
	temp.phen<-merge(temp.totnuc, trait.ds, by="labelcode")
	temp.phen<-dplyr::slice(temp.phen, match(ds0$labelcode, labelcode)) 
	if (trt == "SEX") {
	   phen<-temp.phen %>% dplyr::select(labelcode, age_biopsy, area_name, batch, trt, tot.nuc) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	  # fstr<-paste0("tot.nuc ~ agescale + BATCH + area_name + x + offset(log(tot.nuc))") 
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(BATCH) + as.factor(area_name) + x") 
	}   
	if (trt == "age_biopsy") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, area_name, batch, trt, tot.nuc) %>%
		mutate(BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + x") 
	}   
	if (trt == "T2D") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	  # fstr<-paste0("tot.nuc ~ agescale + BATCH + area_name + x + offset(log(tot.nuc))") 
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + x") 
	}   
	if (!(trt %in% c("SEX", "age_biopsy", "T2D"))) {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc) %>%
		dplyr::filter(complete.cases(.)) %>% 
		dplyr::mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + x") 
	   phen$agescale=scale(phen$age_biopsy)	
	}
	nb<-glm.nb(as.formula(fstr), data=phen)
	#pois<-glm(as.formula(fstr), family="poisson", data=df)
	#ll<-pchisq(2*(logLik(nb) - logLik(pois)), df=1, lower.tail=FALSE)
	#print(ll)
	res<-tidy(nb) %>% filter(substr(term,1,1)=="x")
	# res is for the last variable in the design
	res$trait=trt2
	res$type=type
	res$n=nrow(phen)
	return(res)
}

run_nb_invnorm_bmi<-function(ds0,trt,trt2, type) {
	print(paste0(trt, " " , type ))
	temp<- ds0 %>% dplyr::select(-labelcode) 
	tot.nuc<-rowSums(temp, na.rm=T)
	labelcode<-ds0$labelcode
	temp.totnuc<-cbind.data.frame(labelcode, tot.nuc)
	temp.phen<-merge(temp.totnuc, trait.ds, by="labelcode")
	temp.phen<-dplyr::slice(temp.phen, match(ds0$labelcode, labelcode)) 
	if (trt == "SEX") {
	   phen<-temp.phen %>% dplyr::select(labelcode, age_biopsy, area_name, batch, trt, tot.nuc, bmi) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(BATCH) + as.factor(area_name)  + bmi + x") 
	}   
	if (trt == "age_biopsy") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, area_name, batch, trt, tot.nuc, bmi) %>%
		mutate(BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ as.factor(SEX) + as.factor(BATCH) + as.factor(area_name)  + bmi + x") 
	}   
	if (trt == "T2D") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc, bmi) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	  # fstr<-paste0("tot.nuc ~ agescale + BATCH + area_name + x + offset(log(tot.nuc))") 
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + bmi + x") 
	}   
	if (!(trt %in% c("SEX", "age_biopsy", "T2D"))) {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc, bmi) %>%
		dplyr::filter(complete.cases(.)) %>% 
		dplyr::mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + bmi + x") 
	   phen$agescale=scale(phen$age_biopsy)	
	}
	nb<-glm.nb(as.formula(fstr), data=phen)
	#pois<-glm(as.formula(fstr), family="poisson", data=df)
	#ll<-pchisq(2*(logLik(nb) - logLik(pois)), df=1, lower.tail=FALSE)
	#print(ll)
	res<-tidy(nb) %>% filter(substr(term,1,1)=="x")
	# res is for the last variable in the design
	res$trait=trt2
	res$type=type
	res$n=nrow(phen)
	return(res)
}

run_nb_invnorm_ins<-function(ds0,trt,trt2, type) {
	print(paste0(trt, " " , type ))
	temp<- ds0 %>% dplyr::select(-labelcode) 
	tot.nuc<-rowSums(temp, na.rm=T)
	labelcode<-ds0$labelcode
	temp.totnuc<-cbind.data.frame(labelcode, tot.nuc)
	temp.phen<-merge(temp.totnuc, trait.ds, by="labelcode")
	temp.phen<-dplyr::slice(temp.phen, match(ds0$labelcode, labelcode)) 
	if (trt == "SEX") {
	   phen<-temp.phen %>% dplyr::select(labelcode, age_biopsy, area_name, batch, trt, tot.nuc, S_Insu) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(BATCH) + as.factor(area_name)  + log(S_Insu) + x") 
	}   
	if (trt == "age_biopsy") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, area_name, batch, trt, tot.nuc, S_Insu) %>%
		mutate(BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ as.factor(SEX) + as.factor(BATCH) + as.factor(area_name)  + log(S_Insu) + x") 
	}   
	if (trt == "T2D") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc, S_Insu) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	  # fstr<-paste0("tot.nuc ~ agescale + BATCH + area_name + x + offset(log(tot.nuc))") 
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + log(S_Insu) + x") 
	}   
	if (!(trt %in% c("SEX", "age_biopsy", "T2D"))) {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc, S_Insu) %>%
		dplyr::filter(complete.cases(.)) %>% 
		dplyr::mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + log(S_Insu) + x") 
	   phen$agescale=scale(phen$age_biopsy)	
	}
	nb<-glm.nb(as.formula(fstr), data=phen)
	#pois<-glm(as.formula(fstr), family="poisson", data=df)
	#ll<-pchisq(2*(logLik(nb) - logLik(pois)), df=1, lower.tail=FALSE)
	#print(ll)
	res<-tidy(nb) %>% filter(substr(term,1,1)=="x")
	# res is for the last variable in the design
	res$trait=trt2
	res$type=type
	res$n=nrow(phen)
	return(res)
}

run_nb_invnorm_ogtt<-function(ds0,trt,trt2, type) {
	print(paste0(trt, " " , type ))
	temp<- ds0 %>% dplyr::select(-labelcode) 
	tot.nuc<-rowSums(temp, na.rm=T)
	labelcode<-ds0$labelcode
	temp.totnuc<-cbind.data.frame(labelcode, tot.nuc)
	temp.phen<-merge(temp.totnuc, trait.ds, by="labelcode")
	temp.phen<-dplyr::slice(temp.phen, match(ds0$labelcode, labelcode)) 
	if (trt == "SEX") {
	   phen<-temp.phen %>% dplyr::select(labelcode, age_biopsy, area_name, batch, trt, tot.nuc, ogtt_status_paper) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(BATCH) + as.factor(area_name)  + as.factor(ogtt_status_paper) + x") 
	}   
	if (trt == "age_biopsy") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, area_name, batch, trt, tot.nuc, ogtt_status_paper) %>%
		mutate(BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ as.factor(SEX) + as.factor(BATCH) + as.factor(area_name)  + as.factor(ogtt_status_paper) + x") 
	}   
	if (trt == "T2D") {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc, ogtt_status_paper) %>%
		mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	  # fstr<-paste0("tot.nuc ~ agescale + BATCH + area_name + x + offset(log(tot.nuc))") 
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + as.factor(ogtt_status_paper) + x") 
	}   
	if (!(trt %in% c("SEX", "age_biopsy", "T2D"))) {
	   phen<-temp.phen %>% dplyr::select(labelcode, SEX, age_biopsy, area_name, batch, trt, tot.nuc, ogtt_status_paper) %>%
		dplyr::filter(complete.cases(.)) %>% 
		dplyr::mutate(agescale=scale(age_biopsy), BATCH=paste0("b.",batch)) %>% 
		dplyr::rename(x=trt)
	   phen$x=invnorm(phen$x,1)	
	   fstr<-paste0("tot.nuc ~ agescale + as.factor(SEX) + as.factor(BATCH) + as.factor(area_name) + as.factor(ogtt_status_paper) + x") 
	   phen$agescale=scale(phen$age_biopsy)	
	}
	nb<-glm.nb(as.formula(fstr), data=phen)
	#pois<-glm(as.formula(fstr), family="poisson", data=df)
	#ll<-pchisq(2*(logLik(nb) - logLik(pois)), df=1, lower.tail=FALSE)
	#print(ll)
	res<-tidy(nb) %>% filter(substr(term,1,1)=="x")
	# res is for the last variable in the design
	res$trait=trt2
	res$type=type
	res$n=nrow(phen)
	return(res)
}


dotplot<- function(ds, trtstr, pthresh) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	ds<-ds %>% 
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

	return(ggplot(ds, aes(x=celltype, y=estimate, color=psig)) + 
	geom_point(size=2) + 
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) + 
  	geom_errorbar(aes(ymin=estimate-1.96*(std.error), ymax=estimate+1.96*(std.error))) + 
	scale_colour_manual(values=myColors)  + 
	geom_hline(yintercept=0, colour="gray") + 
  	ggtitle(paste0(trtstr)) + 
	facet_wrap(~modality, strip.position="left", dir="v")
	)
}

dotplot_celltype<-function(ds, cellstr,pthresh,ylim) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	ds<-ds %>% 
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

return(ggplot(ds, aes(x=trait, y=estimate, color=psig)) + 
	geom_point(size=2) + 
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) + 
  	geom_errorbar(aes(ymin=estimate-1.96*(std.error), ymax=estimate+1.96*(std.error))) + 
	scale_colour_manual(values=myColors)  + 
	geom_hline(yintercept=0, colour="gray") + 
  	ggtitle(paste0(cellstr, " NUCLEI")) +
	facet_wrap(~modality, strip.position="left", dir="v")	
	)
}

dotplot_type_rnaatac<-function(ds, types, pthresh,ymin, ymax, sz) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	temp<-ds %>% filter(type %in% all_of(types))  %>%
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

   return(
	ggplot(temp, aes(x=trait, y=estimate, color=psig)) + 
	geom_point(size=2) + 
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5, size=sz)) + 
  	geom_errorbar(aes(ymin=estimate-1.96*(std.error), ymax=estimate+1.96*(std.error))) + 
	ylim(ymin,ymax) + 
	scale_colour_manual(values=myColors)  + 
	theme(legend.position="none") + 
	geom_hline(yintercept=0, colour="gray") + 
	facet_wrap(~type, strip.position="right", dir="v")	
   )
}


pplot_type_rnaatac<-function(ds, types, pthresh,ymin, ymax, sz) {
	myColors = c("black", "blue", "red")
	names(myColors) <- c("No assoc", "Negative assoc", "Positive assoc")

	temp<-ds %>% filter(type %in% all_of(types))  %>%
		mutate(psig=ifelse(padj>=pthresh,"No assoc",
		ifelse(statistic<0,"Negative assoc", "Positive assoc")))

   return(
	ggplot(temp, aes(x=trait, y=-log10(padj), color=psig)) + 
	geom_point(size=2) + 
  	theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5, size=sz)) + 
	ylim(ymin,ymax) + 
	scale_colour_manual(values=myColors)  + 
	theme(legend.position="none") + 
	geom_hline(yintercept=0, colour="gray") + 
	facet_wrap(~type, strip.position="right", dir="v")	
   )
}

 
```

 

Using 279 samples with >=100 RNA and >=100 ATAC nuclei for RNA and both.  
For ATAC, using 281 samples with >=100 ATAC nuclei  


```{r make_phen, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}			
## Make phenotype dataset for DESEq2
ds0 = read_tsv("/net/csgspare2/spare1/muscle_singlenuclei/hg38/fusion_plus_multiome_cluster_info/cluster_info_qc.tsv", guess_max=480000) %>% 
   	rename(labelcode=SNG.1ST) %>% 
	filter(cohort=="FUSION") %>%
	filter(!(labelcode %in% c("12004", "22011", "32071"))) %>% 
	filter(coarse_cluster_name != "Muscle_Fiber_Mixed")
	
batch<-ds0 %>% dplyr::select(labelcode, batch) %>% unique()
heather<- read_csv("data/tissue.csv") 

trait.ds <- heather %>%  mutate(T2D=ifelse(ogtt_status_paper=="T2D",1,ifelse(ogtt_status_paper=="NGT",0,NA)))
#518 samples in heather

trait.ds<-merge(batch, trait.ds, by="labelcode") #284 samples

use1<-readRDS("data/df.both.nuc.Rds")  %>% pull(labelcode)
trait.ds1 <- trait.ds %>% slice(match(use1, labelcode))
saveRDS(trait.ds1, "data/trait.ds.279.Rds") #279 samples

# for atac with 2 more samples...
use2<-readRDS("data/df.atac.nuc.Rds")  %>% pull(labelcode)
trait.ds2 <- trait.ds %>% slice(match(use2, labelcode))
saveRDS(trait.ds2, "data/trait.ds.281.Rds") #281 samples

```


##  SEX only  
Neg Binomial model: total nuclei counts ~ age.scaled + batch + area_name + SEX   
NO SIGNIFICANT ASSOCIATION!!!  

```{r sex_negbinom_output, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE, results='asis'}

df.both.nuc<-readRDS("data/df.both.nuc.Rds")
df.rna.nuc<-readRDS("data/df.rna.nuc.Rds") 
df.atac.nuc<-readRDS("data/df.atac.nuc.Rds") 
trait.ds<-readRDS("data/trait.ds.281.Rds") # atac samples but can be used for both and rna too since match on df.*.nuc samples


traits<-c("SEX")      
traits2<-c("SEX.M") 
  
 i<-1
  
  out<-data.frame()
  res.both<-run_nb_invnorm(df.both.nuc, traits[i], traits2[i], "rna+atac")
   out<-res.both

   res.rna<-run_nb_invnorm(df.rna.nuc, traits[i], traits2[i], "rna") 
   out<-rbind.data.frame(out, res.rna)

   res.atac<-run_nb_invnorm(df.atac.nuc, traits[i], traits2[i], "atac") 
   out<-rbind.data.frame(out, res.atac)
saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_overall_SEXonly_noFDR.Rds")
out<-out %>% dplyr::select(type, trait, estimate, std.error, p.value, n)  
print(kable(out) %>% kable_styling(bootstrap_options=c("striped", "condensed"), full_width=F))
```



```{r make_negbinom_output_invnorm, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}
out<-data.frame()

df.both.nuc<-readRDS("data/df.both.nuc.Rds")
df.rna.nuc<-readRDS("data/df.rna.nuc.Rds") 
df.atac.nuc<-readRDS("data/df.atac.nuc.Rds") 
trait.ds<-readRDS("data/trait.ds.281.Rds") # atac samples but can be used for both and rna too since match on df.*.nuc samples

#65 traits including binary SEX and T2D
  
 traits<-c("SEX", "age_biopsy", "T2D", "S_Insu", "bmi",  "glu_fast_biopsy", "HEIGHT", "WEIGHT", "WAIST", "whr", "HIP", "S_ALAT", "S_hs_CRP", "fS_Kol", "fS_Kol_HDL", "fS_Kol_LDL_c", "fS_Trigly", "GL30", "GL60", "GL120", "S_LipoA1", "S_LipoB", "fS_C_pept", "fS_Krea", "S_GT", "S_Uraat", "fS_C_pept_30", "S_Insu_30", "B_GHb_A1C", "glu_2h_biopsy", "daily_energy_expenditure", "total_daily_physical_activity", "moderate_physical_activity", "strenuous_physical_activity",  "p_insu", "p_insu_30", "p_insu_60", "p_insu_120", "RFM", "sbp", "dbp", "ApoB_A1_ratio", "Ins_AUC_0to30", "Glu_AUC_0to30", "InsSec30", "InsGenIn", "DIo", "CpepGenIn", "HOMA", "matsuda_4pt", "matsuda_3pt", "LTPA_duration_all", "LTPA_duration_cond", "LTPA_duration_noncond", "LTPA_energy_all", "LTPA_energy_cond", "LTPA_energy_noncond", "LTPA_energy_light", "LTPA_energy_moderate", "LTPA_energy_modvig", "LTPA_energy_vigorous", "sleep_biopsy", "totalcw", "sleep_24h", "GL0")      
traits2<-c("SEX.M", "AGE", "T2D", "INSFAST", "BMI", "GLUFAST", "HEIGHT", "WEIGHT", "WAIST", "WHR", "HIP", "ALAT", "CRP", "CHOL", "HDL", "LDL", "TG", "GL30", "GL60", "GL120", "APOA1", "APOB", "CPEP", "CREAT", "GT", "URIC", "CPEP30", "S_INS30", "HBA1C", "GLU2H", "ENERGY", "PA_TOT", "PA_MOD", "PA_STREN", "INS0", "INS30", "INS60", "INS120", "RFM", "SBP", "DBP", "APOB.A1.RATIO", "INS_AUC_0_30", "GLU_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG", "SLEEP", "ALCOHOL", "SLEEP24H", "GL0")   
 
   out<-data.frame()
   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.both<-run_nb_invnorm(df.both.nuc, traits[i], traits2[i], "rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   #make B-H adjustment across traits
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-tempout


   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.rna<-run_nb_invnorm(df.rna.nuc, traits[i], traits2[i], "rna") 
	if (i==1) {tempout=res.rna} else {tempout=rbind.data.frame(tempout,res.rna)}
   }
   #make B-H adjustment across traits
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)

   tempout<-data.frame()
   for (i in 1:length(traits)) {
	res.atac<-run_nb_invnorm(df.atac.nuc, traits[i], traits2[i],"atac") 
	if (i==1) {tempout=res.atac} else {tempout=rbind.data.frame(tempout,res.atac)}
	}
   #make B-H adjustment across traits
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)
saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_overall.Rds")
```

##  Neg Binomial model: nuclei counts ~ age.scaled + sex + batch + area_name + trait.invnorm    
Dot plots are the neg binomial estimates +/- 1.96*SE   
Red/blue 'significant' represents B-H q<.05, calculated across 65 traits within each type (rna, atac, rna+atac)

```{r nucleus_type_traits_invnorm, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates or p-values for each invnorm trait, per nucleus type"}

types1<-c("rna", "atac", "rna+atac")

df<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_overall.Rds")
df$trait<-factor(df$trait, levels=c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "INSFAST", "INS0", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE"))
dot.base1<-dotplot_type_rnaatac(df,types1,0.05,-.25,0.25, 6)
dot.base<-dotplot_type_rnaatac(df,types1,0.05,-.25,0.3, 4) #for comparison plots in code chunk nucleus_type_traits_invnorm_bmi
pv.base<-pplot_type_rnaatac(df,types1,0.05,0,4, 6)
print(dot.base1)
print(pv.base)
```


##  Adjusting for BMI  
Neg Binomial model: nuclei counts ~ age.scaled + sex + batch + area_name + BMI + trait.invnorm    
```{r make_negbinom_output_invnorm_bmi, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}
out<-data.frame()

df.both.nuc<-readRDS("data/df.both.nuc.Rds")
df.rna.nuc<-readRDS("data/df.rna.nuc.Rds") 
df.atac.nuc<-readRDS("data/df.atac.nuc.Rds") 
trait.ds<-readRDS("data/trait.ds.281.Rds") # atac samples but can be used for both and rna too since match on df.*.nuc samples
traits.nobmi<-traits[traits !="bmi"]
traits2.nobmi<-traits2[traits2 !="BMI"]
  
   out<-data.frame()
   tempout<-data.frame()
   for (i in 1:length(traits.nobmi)) {
	res.both<-run_nb_invnorm_bmi(df.both.nuc, traits.nobmi[i], traits2.nobmi[i], "rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   #make B-H adjustment across traits
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-tempout


   tempout<-data.frame()
   for (i in 1:length(traits.nobmi)) {
	res.rna<-run_nb_invnorm_bmi(df.rna.nuc, traits.nobmi[i], traits2.nobmi[i], "rna") 
	if (i==1) {tempout=res.rna} else {tempout=rbind.data.frame(tempout,res.rna)}
   }
   #make B-H adjustment across traits
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)

   tempout<-data.frame()
   for (i in 1:length(traits.nobmi)) {
	res.atac<-run_nb_invnorm_bmi(df.atac.nuc, traits.nobmi[i], traits2.nobmi[i], "atac") 
	if (i==1) {tempout=res.atac} else {tempout=rbind.data.frame(tempout,res.atac)}
	}
   #make B-H adjustment across traits
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)
saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_overall_adjbmi.Rds")
```

```{r nucleus_type_traits_invnorm_bmi, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="nb estimates for each invnorm trait, per nucleus type, adjusted for bmi"}

types1<-c("rna", "atac", "rna+atac")

df<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_overall_adjbmi.Rds")
df$trait<-factor(df$trait, levels=c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "INSFAST", "INS0", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE"))
dot.bmi<-dotplot_type_rnaatac(df,types1,0.05,-.25,0.3, 4)
print(grid.arrange(dot.base, dot.bmi, nrow=1, ncol=2))
#pv.bmi<-pplot_type_rnaatac(df,types1,0.05,0,4, 4)
#print(grid.arrange(pv.base, pv.bmi, nrow=1, ncol=2))
```




##  Adjusting for fasting insulin kills associations with total nuclei    
Neg Binomial model: nuclei counts ~ age.scaled + sex + batch + area_name + log(S_Insu) + trait.invnorm   
```{r make_negbinom_output_invnorm_ins, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}
out<-data.frame()

df.both.nuc<-readRDS("data/df.both.nuc.Rds")
df.rna.nuc<-readRDS("data/df.rna.nuc.Rds") 
df.atac.nuc<-readRDS("data/df.atac.nuc.Rds") 
trait.ds<-readRDS("data/trait.ds.281.Rds")
traits.noins<-traits[traits !="S_Insu"]
traits2.noins<-traits2[traits2 !="INSFAST_SERUM"]
  
   out<-data.frame()
   tempout<-data.frame()
   for (i in 1:length(traits.noins)) {
	res.both<-run_nb_invnorm_ins(df.both.nuc, traits.noins[i], traits2.noins[i], "rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-tempout


   tempout<-data.frame()
   for (i in 1:length(traits.noins)) {
	res.rna<-run_nb_invnorm_ins(df.rna.nuc, traits.noins[i], traits2.noins[i], "rna") 
	if (i==1) {tempout=res.rna} else {tempout=rbind.data.frame(tempout,res.rna)}
   }
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)

   tempout<-data.frame()
   for (i in 1:length(traits.noins)) {
	res.atac<-run_nb_invnorm_ins(df.atac.nuc, traits.noins[i], traits2.noins[i], "atac") 
	if (i==1) {tempout=res.atac} else {tempout=rbind.data.frame(tempout,res.atac)}
	}
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)
saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_overall_adjins.Rds")
```

```{r nucleus_type_traits_invnorm_ins, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="left: all samples, base model, right: all samples, with log(fasting insulin) covariate "}
types1<-c("rna", "atac", "rna+atac")


df<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_overall_adjins.Rds")
df$trait<-factor(df$trait, levels=c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "INSFAST", "INS0", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE"))
dot.ins<- dotplot_type_rnaatac(df,types1,0.05,-.7,0.6, 4)
pv.ins<-pplot_type_rnaatac(df,types1,0.05,0,4, 4)
print(grid.arrange(dot.base, dot.ins, nrow=1, ncol=2))
#print(grid.arrange(pv.base, pv.ins, nrow=1, ncol=2))

```






##  Adjusting for ogtt status :  matsuda index becomes significant in positive direction
Neg Binomial model: nuclei counts ~ age.scaled + sex + batch + area_name + as.factor(ogtt_status_paper) + trait.invnorm   
```{r make_negbinom_output_invnorm_ogtt, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=TRUE}

df.both.nuc<-readRDS("data/df.both.nuc.Rds")
df.rna.nuc<-readRDS("data/df.rna.nuc.Rds") 
df.atac.nuc<-readRDS("data/df.atac.nuc.Rds") 
trait.ds<-readRDS("data/trait.ds.281.Rds")
traits.no.ogtt<-traits[traits !="T2D"]
traits2.no.ogtt<-traits2[traits2 !="T2D"]
  
   out<-data.frame()
   tempout<-data.frame()
   for (i in 1:length(traits.no.ogtt)) {
	res.both<-run_nb_invnorm_ogtt(df.both.nuc, traits.no.ogtt[i], traits2.no.ogtt[i], "rna+atac")
	if (i==1) {tempout=res.both} else {tempout=rbind.data.frame(tempout,res.both)}
   }
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-tempout


   tempout<-data.frame()
   for (i in 1:length(traits.no.ogtt)) {
	res.rna<-run_nb_invnorm_ogtt(df.rna.nuc, traits.no.ogtt[i], traits2.no.ogtt[i], "rna") 
	if (i==1) {tempout=res.rna} else {tempout=rbind.data.frame(tempout,res.rna)}
   }
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)

   tempout<-data.frame()
   for (i in 1:length(traits.no.ogtt)) {
	res.atac<-run_nb_invnorm_ogtt(df.atac.nuc, traits.no.ogtt[i], traits2.no.ogtt[i], "atac") 
	if (i==1) {tempout=res.atac} else {tempout=rbind.data.frame(tempout,res.atac)}
	}
   #make B-H adjustment across traits, within cell type
   tempout<- tempout %>% mutate(padj=p.adjust(p.value, method="BH"))
   out<-rbind.data.frame(out, tempout)
saveRDS(out, "output/nuclei_nb/nb_nuclei_invnorm_overall_adjogtt.Rds")
```

```{r nucleus_type_traits_invnorm_ogtt, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, eval=TRUE, results='hide', fig.cap="left: all samples, base model, right: all samples, with ogtt_status_paper covariate "}
types1<-c("rna", "atac", "rna+atac")


df<-readRDS("output/nuclei_nb/nb_nuclei_invnorm_overall_adjogtt.Rds")
df$trait<-factor(df$trait, levels=c("HEIGHT", "BMI", "HIP", "RFM", "WAIST", "WEIGHT", "WHR", "CHOL", "HDL", "LDL", "TG", "APOA1", "APOB", "APOB.A1.RATIO", "GLUFAST", "GL0", "GL30", "GL60", "GL120", "GLU_AUC_0_30", "GLU2H",  "HBA1C", "T2D", "INSFAST", "INS0", "S_INS30", "INS30", "INS60", "INS120", "CPEP", "CPEP30", "INS_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "CRP", "CREAT", "GT", "ALAT", "URIC", "SBP", "DBP", "ENERGY",  "PA_TOT", "PA_MOD", "PA_STREN", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG",  "SLEEP", "SLEEP24H", "ALCOHOL", "SEX.M", "AGE"))
dot.ogtt<- dotplot_type_rnaatac(df,types1,0.05,-.25,0.3, 4)
#pv.ogtt<-pplot_type_rnaatac(df,types1,0.05,0,4, 4)
print(grid.arrange(dot.base, dot.ogtt, nrow=1, ncol=2))

```


