---
title: "DESeq2 negative binomial regression using gene count data in bulk,in hg38."
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, dev=c('png'), include=F}
knitr::opts_chunk$set(dev = c('png'))
knitr::opts_chunk$set(autodep = TRUE)

```
Running all traits including sex.  Restricting to bulk genes with 5 counts in at least 25% of the samples.  Limiting to Type_1 samples.

Please note this was run in slurm by converting this to code/run_deseq_RNA_bulk_alltraits.R 
% sbatch --array=1-66 Run_deseq_RNA_bulk_alltraits.sh
deseq.RNA.bulk.alltraits.jobs made from code/make_deseq_RNA_bulk_alltraits_jobs.sh

```{r include=F}
.libPaths("/net/snowwhite/home/aujackso/R/x86_64-pc-linux-gnu-library/4.0");
library(dplyr) 
library(readr)
library(ggplot2)
library(DESeq2) 
library(edgeR)
#library(RColorBrewer)
#library(grid)
#library(gridExtra)
library(knitr)
library(stringr)
options(scipen=1, digits=5)
options(nwarnings=10000)

invnorm = function(x, seed) {
  set.seed(seed)
  qnorm((rank(x, na.last="keep", ties.method="random") - 0.5) / sum(!is.na(x)))
}
# h.traits doesn't include T2D or traits previously in bulk phenotype dataset from 6.1_bulk_pbulk_deseq_hg38.Rmd
h.traits<-c("S_Insu", "glu_fast_biopsy", "HEIGHT", "WEIGHT", "WAIST", "whr", "HIP", "S_ALAT", "S_hs_CRP", "fS_Kol", "fS_Kol_HDL", "fS_Kol_LDL_c", "fS_Trigly", "GL0", "GL30", "GL60", "GL120", "S_LipoA1", "S_LipoB", "fS_C_pept", "fS_Krea", "S_GT", "S_Uraat", "fS_C_pept_30", "S_Insu_30", "B_GHb_A1C", "glu_2h_biopsy", "daily_energy_expenditure", "total_daily_physical_activity", "moderate_physical_activity", "strenuous_physical_activity",  "p_insu", "p_insu_30", "p_insu_60", "p_insu_120", "RFM", "sbp", "dbp", "ApoB_A1_ratio", "Ins_AUC_0to30", "Glu_AUC_0to30", "InsSec30", "InsGenIn", "DIo", "CpepGenIn", "HOMA", "matsuda_4pt", "matsuda_3pt", "LTPA_duration_all", "LTPA_duration_cond", "LTPA_duration_noncond", "LTPA_energy_all", "LTPA_energy_cond", "LTPA_energy_noncond", "LTPA_energy_light", "LTPA_energy_moderate", "LTPA_energy_modvig", "LTPA_energy_vigorous", "sleep_biopsy", "totalcw", "sleep_24h","GL0")      


traits<-c("sex", "age_at_biopsy", "T2D", "S_Insu", "bmi",  "glu_fast_biopsy", "HEIGHT", "WEIGHT", "WAIST", "whr", "HIP", "S_ALAT", "S_hs_CRP", "fS_Kol", "fS_Kol_HDL", "fS_Kol_LDL_c", "fS_Trigly", "GL0", "GL30", "GL60", "GL120", "S_LipoA1", "S_LipoB", "fS_C_pept", "fS_Krea", "S_GT", "S_Uraat", "fS_C_pept_30", "S_Insu_30", "B_GHb_A1C", "glu_2h_biopsy", "daily_energy_expenditure", "total_daily_physical_activity", "moderate_physical_activity", "strenuous_physical_activity",  "p_insu", "p_insu_30", "p_insu_60", "p_insu_120", "RFM", "sbp", "dbp", "ApoB_A1_ratio", "Ins_AUC_0to30", "Glu_AUC_0to30", "InsSec30", "InsGenIn", "DIo", "CpepGenIn", "HOMA", "matsuda_4pt", "matsuda_3pt", "LTPA_duration_all", "LTPA_duration_cond", "LTPA_duration_noncond", "LTPA_energy_all", "LTPA_energy_cond", "LTPA_energy_noncond", "LTPA_energy_light", "LTPA_energy_moderate", "LTPA_energy_modvig", "LTPA_energy_vigorous", "sleep_biopsy", "totalcw", "sleep_24h","GL0")      
traits2<-c("SEX.M", "AGE", "T2D", "INSFAST", "BMI", "GLUFAST", "HEIGHT", "WEIGHT", "WAIST", "WHR", "HIP", "ALAT", "CRP", "CHOL", "HDL", "LDL", "TG", "GL0", "GL30", "GL60", "GL120", "APOA1", "APOB", "CPEP", "CREAT", "GT", "URIC", "CPEP30", "S_INS30", "HBA1C", "GLU2H", "ENERGY", "PA_TOT", "PA_MOD", "PA_STREN", "INS0", "INS30", "INS60", "INS120", "RFM", "SBP", "DBP", "APOB.A1.RATIO", "INS_AUC_0_30", "GLU_AUC_0_30", "INSSEC30", "INSGENIN", "DI", "CPEPGENIN", "HOMA", "MATSUDA4", "MATSUDA3", "LTPA.DUR.ALL", "LTPA.DUR.COND", "LTPA.DUR.NONCOND", "LTPA.EN.ALL", "LTPA.EN.COND", "LTPA.EN.NONCOND", "LTPA.EN.LIGHT", "LTPA.EN.MOD", "LTPA.EN.MODVIG", "LTPA.EN.VIG", "SLEEP", "ALCOHOL", "SLEEP24H","GL0") 
 #66 traits 


```
Run bulk using bulk DESeq. For bulk use typical bulk base covariates, base + 3 fiber type proportions, base + snRNA cell type proportions.  
Use older DESeq parameters (not for single cell).       
Restrict to samples in snRNA since using cell type proportions as covariates
Limit genes to those with at least 5 counts in at least 25% of samples

Using datasets created in 6.1_bulk_pbulk_deseq_hg38.Rmd

```{r read_data}

p.bulk<-readRDS("data/phen.bulk.Rds") %>% 
		replace(is.na(.), 0) %>% # 4 macrophage and 1 adipocyte proportion are NA so set to 0

		mutate(T2D=ifelse(ogtt=="T2D",1,ifelse(ogtt=="NGT",0,NA))) #115 T2D NAs
		
heather<-read_csv("data/tissue.csv") %>% select(labelcode, all_of(h.traits))

p.bulk<-merge(p.bulk, heather, by="labelcode")
  #lots of traits have missing values

# 22323 genes filtered for having at least 5 counts in at least 25% of the 268 samples in the bulk dataset after overlapping the samples with snRNA (pseudobulk)
#  Note that I am NOT dropping genes later based on 5 counts in at least 25% of samples with phenotypes!
bulk.filt<-readRDS("data/ct.bulk.filt.Rds") #22323 genes, 268 samples
annot.sn<-readRDS("data/annot.sn.Rds")
```


## Base covariates model FILTERING FOR LOW COUNTS     
bulk counts ~  SEX + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + TRAIT  
```{r run_bulk_deseq_base_filter}

   for (j in 1:length(traits)) {
	
	eds<-bulk.filt

	if (traits[j]=="SEX") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median) 

		fstr<-"~  AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT +  as.factor(sex)" 
	}
	if (traits[j]=="T2D") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, T2D) %>%
		dplyr::filter(complete.cases(.)) 
		
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT +  T2D" 
	}
	if (traits[j]=="AGE") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median) 
		
		fstr<-"~  as.factor(sex) + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + AGE" 
	}
	if (!(traits[j] %in% c("SEX", "AGE", "T2D"))) {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, traits[j]) %>%
		dplyr::filter(complete.cases(.)) 
		pds$x<- pds %>% pull(traits[j])
		pds$x<-invnorm(pds$x, j)
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + x" 
	}

#inverse normalize covariates based on available trait
pds<-pds %>% 
	mutate(AGE=invnorm(age_at_biopsy,1), RIN=invnorm(rin,1), TIN=invnorm(tin_median,1), GC=invnorm(GC_byRead_Mean,1), INSERT=invnorm(InsertSize_Median,1)) 
samps<-pds %>% pull(labelcode) #to select gene columns
eds<-as.data.frame(eds) %>% dplyr::select(all_of(samps)) %>% as.matrix()	

#  Using original DESeq bulk settings
	dds <- DESeqDataSetFromMatrix(countData = eds,
            colData = pds, design = as.formula(fstr)) 
	#dds<-scran::computeSumFactors(dds)   
	#dds<-DESeq(dds, test="LRT",reduced=as.formula(redstr), useT=TRUE, minmu=1e-6, minReplicatesForReplace=Inf, fitType="local")  
        #dds<-DESeq(dds, test="Wald", fitType="parametric",parallel=TRUE) 
	dds<-estimateSizeFactors(dds)
	dds<-estimateDispersions(dds) 
        dds<-nbinomWaldTest(dds, maxit=1000)
	stats<-mcols(dds) %>% as.data.frame() 
	stats$gene=rownames(stats)
	converge<-stats %>% dplyr::select(gene, betaConv)
	
	res = as(results(dds), "data.frame") 
	res$gene=rownames(res)
	res$data="bulk.base"
	res$trait=traits2[j]
	res$n=nrow(pds)
	res<-merge(res,converge,by="gene")
	
res1<-merge(res, annot.sn, by="gene") %>% arrange(pvalue)
write.table(format(res1, digits=8), 
file=paste0("~/snRNAsnATAC_paper1/output/DESeq.RNA/bulk_alltraits/hg38.filter22323.bulk.base.",traits2[j],".results.tab"), sep="\t", quote=F, col.names=T, row.names=F)
}

```


## Base + fiber proxy covariates model   - Filter for low counts    
bulk counts ~  SEX + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + trait      
Use Type 1 as reference fiber type proxy    
```{r run_bulk_deseq_basefiber_filter}

   for (j in 1:length(traits)) {
	
	eds<-bulk.filt

	if (traits[j]=="SEX") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, type2A.fiber.proxy, type2X.fiber.proxy) 
		
		fstr<-"~  AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + as.factor(sex)" 
	}
	if (traits[j]=="T2D") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, type2A.fiber.proxy, type2X.fiber.proxy, T2D) %>%
		dplyr::filter(complete.cases(.)) 
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + T2D" 
	}
	if (traits[j]=="AGE") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, type2A.fiber.proxy, type2X.fiber.proxy) 
		
		fstr<-"~  as.factor(sex) + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + AGE" 
	}
	if (!(traits[j] %in% c("SEX", "AGE", "T2D"))) {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, traits[j], type2A.fiber.proxy, type2X.fiber.proxy) %>%
		dplyr::filter(complete.cases(.)) 
		pds$x<- pds %>% pull(traits[j])
		pds$x<-invnorm(pds$x, j)
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + x" 
	}

#inverse normalize covariates based on available phenotype
pds <- pds %>% 
	mutate(AGE=invnorm(age_at_biopsy,1), RIN=invnorm(rin,1), TIN=invnorm(tin_median,1), GC=invnorm(GC_byRead_Mean,1), INSERT=invnorm(InsertSize_Median,1), PROXY.2A=invnorm(type2A.fiber.proxy,2), PROXY.2X=invnorm(type2X.fiber.proxy,2))

samps<-pds %>% pull(labelcode) #to select gene columns
eds<-as.data.frame(eds) %>% dplyr::select(all_of(samps)) %>% as.matrix()	

#  Using original DESeq bulk settings
	dds <- DESeqDataSetFromMatrix(countData = eds,
            colData = pds, design = as.formula(fstr)) 
	dds<-estimateSizeFactors(dds)
	dds<-estimateDispersions(dds) 
        dds<-nbinomWaldTest(dds, maxit=1000)
	stats<-mcols(dds) %>% as.data.frame() 
	stats$gene=rownames(stats)
	converge<-stats %>% dplyr::select(gene, betaConv)
	
	res = as(results(dds), "data.frame") 
	res$gene=rownames(res)
	res$data="bulk.basefiber"
	res$trait=traits2[j]
	res$n=nrow(pds)
	res<-merge(res,converge,by="gene")
	
res1<-merge(res, annot.sn, by="gene") %>% arrange(pvalue)
write.table(format(res1, digits=8), 
file=paste0("~/snRNAsnATAC_paper1/output/DESeq.RNA/bulk_alltraits/hg38.filter22323.bulk.basefiber.",traits2[j],".results.tab"), sep="\t", quote=F, col.names=T, row.names=F)
}

```



## Base + fiber proxy + cell type proportion covariates model - Filter for low counts     
bulk counts ~  SEX + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + TYPE2A + TYPE2X + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + trait 
Use Type 1 as reference fiber type proxy and reference cell type.     

```{r run_bulk_deseq_basefibercelltype_filter}

   for (j in 1:length(traits)) {
	
	eds<-bulk.filt

	if (traits[j]=="SEX") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, type2A.fiber.proxy, type2X.fiber.proxy, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) 
		
		fstr<-"~  AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + as.factor(sex)" 
	}
	if (traits[j]=="T2D") {
	pds <- p.bulk %>% 
		dplyr::select(T2D, sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, type2A.fiber.proxy, type2X.fiber.proxy, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) %>%
		dplyr::filter(complete.cases(.)) 
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + T2D" 
	}
	if (traits[j]=="AGE") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, type2A.fiber.proxy, type2X.fiber.proxy, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) 
		
		fstr<-"~  as.factor(sex) + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + AGE" 
	}
	if (!(traits[j] %in% c("SEX", "AGE", "T2D"))) {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, traits[j], type2A.fiber.proxy, type2X.fiber.proxy, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) %>%
		dplyr::filter(complete.cases(.)) 
		pds$x<- pds %>% pull(traits[j])
		pds$x<-invnorm(pds$x, j)
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + PROXY.2A + PROXY.2X + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + x" 
	}

#inverse normalize covariates based on available phenotype
pds <- pds %>% 
	mutate(AGE=invnorm(age_at_biopsy,1), RIN=invnorm(rin,1), TIN=invnorm(tin_median,1), GC=invnorm(GC_byRead_Mean,1), INSERT=invnorm(InsertSize_Median,1), PROXY.2A=invnorm(type2A.fiber.proxy,2), PROXY.2X=invnorm(type2X.fiber.proxy,2), TYPE2A = invnorm(Type_2a,3), TYPE2X=invnorm(Type_2x,3), ADI=invnorm(Adipocyte,3), ENDO=invnorm(Endothelial,3), MACRO=invnorm(Macrophage,3), MSC=invnorm(Mesenchymal_Stem_Cell,3), NMJ=invnorm(Neuromuscular_junction,3), NEUR=invnorm(Neuronal,3), SATEL=invnorm(Satellite_Cell,3), SM=invnorm(Smooth_Muscle,3), TC=invnorm(T_cell,3))

samps<-pds %>% pull(labelcode) #to select gene columns
eds<-as.data.frame(eds) %>% dplyr::select(all_of(samps)) %>% as.matrix()	

#  Using original DESeq bulk settings
	dds <- DESeqDataSetFromMatrix(countData = eds,
            colData = pds, design = as.formula(fstr)) 
	dds<-estimateSizeFactors(dds)
	dds<-estimateDispersions(dds) 
        dds<-nbinomWaldTest(dds, maxit=1000)
	stats<-mcols(dds) %>% as.data.frame() 
	stats$gene=rownames(stats)
	converge<-stats %>% dplyr::select(gene, betaConv)
	
	res = as(results(dds), "data.frame") 
	res$gene=rownames(res)
	res$data="bulk.basefibercelltype"
	res$trait=traits2[j]
	res$n=nrow(pds)
	res<-merge(res,converge,by="gene")
	
res1<-merge(res, annot.sn, by="gene") %>% arrange(pvalue)
write.table(format(res1, digits=8), 
file=paste0("~/snRNAsnATAC_paper1/output/DESeq.RNA/bulk_alltraits/hg38.filter22323.bulk.basefibercelltype.",traits2[j],".results.tab"), sep="\t", quote=F, col.names=T, row.names=F)
}

```


## Base + cell type proportion covariates model - Filter for low counts     
bulk counts ~  SEX + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + trait    
Use Type 1 as reference cell type.     


```{r run_bulk_deseq_basecelltype_filter}


   for (j in 1:length(traits)) {
	
	eds<-bulk.filt

	if (traits[j]=="SEX") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) 
		
		fstr<-"~  AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + as.factor(sex)" 
	}
	if (traits[j]=="T2D") {
	pds <- p.bulk %>% 
		dplyr::select(T2D, sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) %>%
		dplyr::filter(complete.cases(.)) 
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + T2D" 
	}
	if (traits[j]=="AGE") {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) 
		
		fstr<-"~  as.factor(sex) + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + AGE" 
	}
	if (!(traits[j] %in% c("SEX", "AGE", "T2D"))) {
	pds <- p.bulk %>% 
		dplyr::select(sex, labelcode, age_at_biopsy, mrna_batch, area_name, rin, tin_median, GC_byRead_Mean, InsertSize_Median, traits[j], Type_2a, Type_2x, Adipocyte, Endothelial, Macrophage, Mesenchymal_Stem_Cell, Neuromuscular_junction, Neuronal, Satellite_Cell, Smooth_Muscle, T_cell) %>%
		dplyr::filter(complete.cases(.)) 
		pds$x<- pds %>% pull(traits[j])
		pds$x<-invnorm(pds$x, j)
		fstr<-"~  as.factor(sex) + AGE + as.factor(mrna_batch) + RIN + TIN + as.factor(area_name) + GC + INSERT + ADI + ENDO + MACRO + MSC + NMJ + NEUR + SATEL + SM + TC + x" 
	}

#inverse normalize covariates based on available phenotype
pds <- pds %>% 
	mutate(AGE=invnorm(age_at_biopsy,1), RIN=invnorm(rin,1), TIN=invnorm(tin_median,1), GC=invnorm(GC_byRead_Mean,1), INSERT=invnorm(InsertSize_Median,1), TYPE2A = invnorm(Type_2a,3), TYPE2X=invnorm(Type_2x,3), ADI=invnorm(Adipocyte,3), ENDO=invnorm(Endothelial,3), MACRO=invnorm(Macrophage,3), MSC=invnorm(Mesenchymal_Stem_Cell,3), NMJ=invnorm(Neuromuscular_junction,3), NEUR=invnorm(Neuronal,3), SATEL=invnorm(Satellite_Cell,3), SM=invnorm(Smooth_Muscle,3), TC=invnorm(T_cell,3))

samps<-pds %>% pull(labelcode) #to select gene columns
eds<-as.data.frame(eds) %>% dplyr::select(all_of(samps)) %>% as.matrix()	

#  Using original DESeq bulk settings
	dds <- DESeqDataSetFromMatrix(countData = eds,
            colData = pds, design = as.formula(fstr)) 
	dds<-estimateSizeFactors(dds)
	dds<-estimateDispersions(dds) 
        dds<-nbinomWaldTest(dds, maxit=1000)
	stats<-mcols(dds) %>% as.data.frame() 
	stats$gene=rownames(stats)
	converge<-stats %>% dplyr::select(gene, betaConv)
	
	res = as(results(dds), "data.frame") 
	res$gene=rownames(res)
	res$data="bulk.basecelltype"
	res$trait=traits2[j]
	res$n=nrow(pds)
	res<-merge(res,converge,by="gene")
	
res1<-merge(res, annot.sn, by="gene") %>% arrange(pvalue)
write.table(format(res1, digits=8), 
file=paste0("~/snRNAsnATAC_paper1/output/DESeq.RNA/bulk_alltraits/hg38.filter22323.bulk.basecelltype.",traits2[j],".results.tab"), sep="\t", quote=F, col.names=T, row.names=F)
}
```




