---
title: "DESeq2 negative binomial regression using gene count data in GTEX bulk,in hg38."
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, dev=c('png'), include=F}
knitr::opts_chunk$set(dev = c('png'))
knitr::opts_chunk$set(autodep = TRUE)

```


```{r include=F}
.libPaths("/net/snowwhite/home/aujackso/R/x86_64-pc-linux-gnu-library/4.0");
library(dplyr) 
library(readr)
library(ggplot2)
library(DESeq2) 
library(edgeR)
#library(RColorBrewer)
#library(grid)
#library(gridExtra)
library(knitr)
library(stringr)
options(scipen=1, digits=5)
options(nwarnings=10000)

invnorm = function(x, seed) {
  set.seed(seed)
  qnorm((rank(x, na.last="keep", ties.method="random") - 0.5) / sum(!is.na(x)))
}


#  Argh need to update R https://bioconductor.org/packages/release/bioc/html/IRanges.html
```
Run bulk using bulk DESeq. using original DESeq parameters (not for single cell).       
Restrict to samples in snRNA since using cell type proportions as covariates
Limit genes to those with at least 5 counts in at least 25% of samples

=========   FILTER FOR LOW COUNTS =====================================
#  Drop genes missing 5 counts in at least 25% of samples
```{r prep_ds_filter, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, eval=FALSE}

attrib0<-read_tsv("data/gtex/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt") %>% filter(SMTS=="Muscle") %>% mutate(SUBJID=word(SAMPID, start=1,end=2, sep=fixed("-"))) %>% select(SUBJID, SMRIN, SMNABTCH, SMCENTER, SAMPID)
length(unique(attrib0$SUBJID))
#attrib has 1132 SAMPIDs but only 854 uniq SUBJIDs initially

phen<-read_tsv("data/gtex/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt") %>% select(SUBJID, SEX, AGE, DTHHRDY) 
#980 subjects


ct.gtex <-read_tsv("data/gtex/gene_reads_muscle_skeletal.gct", skip=2)

# counts has 56200 genes for 803 SAMPIDs
who.gtex<-colnames(ct.gtex)[4:ncol(ct.gtex)]
# limit attrib to the 803
attrib<- attrib0 %>% dplyr::slice(match(who.gtex, SAMPID))
length(unique(attrib$SUBJID))

gtex_phen0<-left_join(attrib, phen, by="SUBJID")
all.equal(gtex_phen0$SAMPID, who.gtex)


table(gtex_phen0$SEX, useNA='ifany')
#  1   2 
#543 260
table(gtex_phen0$AGE, useNA='ifany')
#20-29 30-39 40-49 50-59 60-69 70-79 
#   67    65   124   255   264    28 
table(gtex_phen0$DTHHRDY, useNA='ifany')
#   0    1    2    3    4 <NA> 
# 424   31  203   46   87   12 
summary(gtex_phen0$SMRIN)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   5.50    7.70    8.00    7.94    8.40    9.90 

length(unique(gtex_phen0$SMNABTCH))
#  344 levels... 1-7 samples per batch

table(gtex_phen0$SMCENTER, useNA='ifany')
#  B1   C1   D1 <NA> 
# 548  251    3    1 


gtex_phen <- gtex_phen0 %>% filter(complete.cases(.))
#  But only 790 with complete covariates - 12 were missing DTHHRDY death classification and 1 missing CENTER

table(gtex_phen$SEX, useNA='ifany')
#  1   2 
#534 256
table(gtex_phen$AGE, useNA='ifany')
#20-29 30-39 40-49 50-59 60-69 70-79 
#   67    64   122   250   261    26 
table(gtex_phen$DTHHRDY, useNA='ifany')
#  0   1   2   3   4 
#423  31 203  46  87 
summary(gtex_phen$SMRIN)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#   5.50    7.70    8.00    7.94    8.40    9.90 

length(unique(gtex_phen$SMNABTCH))
#  338 levels... 1-7 samples per batch

table(gtex_phen$SMCENTER, useNA='ifany')
# B1  C1  D1 
#538 249   3 


gtex_phen<-gtex_phen %>% dplyr::rename(sexn=SEX) %>% mutate(SEX=ifelse(sexn==1,"M","F"))

saveRDS(gtex_phen, "data/gtex.phen.Rds")



exp0<-ct.gtex %>% select(gtex_phen$SAMPID) %>% as.matrix()
#id is number 0+, Name is Ensembl gene ID, Description is gene symbol
row.names(exp0)<-ct.gtex$Name

gtex.exp<-DGEList(counts=exp0) #56200
gtex.sample_size <- ncol(exp0) #790
gene.keep<-rowSums(gtex.exp$counts >=5) >= round(0.25*gtex.sample_size) #T/F
gtex.filter<-gtex.exp[gene.keep,] 
gtex.genes.filter<-rownames(gtex.filter) 

gtex.filt<-as.matrix(gtex.filter[[1]])
saveRDS(gtex.filt, "data/gtex.filt.Rds")
#22211 genes, 790 samples

```

		

## Base covariates model FILTERING FOR LOW COUNTS     
Count ~ Sex + Age(AGE)  + Hardy Scale(DTHHRDY) (these three from phenotype file) + SMRIN + SMNABTCH + SMCENTER (these three from the attributes file)

```{r run_bulk_deseq_base_filter}
	
	
eds<-gtex.filt

pds <- gtex_phen 
#fstr<-"~  as.factor(AGE) + as.factor(DTHHRDY) + RIN + as.factor(SMNABTCH) + as.factor(SMCENTER) +  as.factor(SEX)" 
#drop batch since get 'Model matrix not full rank' errors from DESeq2
fstr<-"~  as.factor(AGE) + as.factor(DTHHRDY) + RIN + as.factor(SMCENTER) +  as.factor(SEX)" 

#inverse normalize covariates based on available trait
pds<-pds %>% 
	mutate(RIN=invnorm(SMRIN,1)) 
all.equal(colnames(gtex.filt), pds$SAMPID)

#  Using original DESeq bulk settings
dds <- DESeqDataSetFromMatrix(countData = eds,
       colData = pds, design = as.formula(fstr)) 
	#dds<-scran::computeSumFactors(dds)   
	#dds<-DESeq(dds, test="LRT",reduced=as.formula(redstr), useT=TRUE, minmu=1e-6, minReplicatesForReplace=Inf, fitType="local")  
        #dds<-DESeq(dds, test="Wald", fitType="parametric",parallel=TRUE) 
dds<-estimateSizeFactors(dds)
dds<-estimateDispersions(dds) 
dds<-nbinomWaldTest(dds, maxit=1000)
stats<-mcols(dds) %>% as.data.frame() 
stats$gene=rownames(stats)
converge<-stats %>% dplyr::select(gene, betaConv)
	
res = as(results(dds), "data.frame") 
res$gene=rownames(res)
res$data="gtex"
res$trait="SEX"
res$n=nrow(pds)
res<-merge(res,converge,by="gene")
	 
	# annotation from Arushi for our hg38 muscle data
	annot.sn0<-read_tsv("data/gencode.v30.gene_lengths_tss.tsv") %>% 
		mutate(chr=gsub("chr","",chrom)) %>% 
		dplyr::rename(start=gene_start, end=gene_end, symbol=gene_name, strand=gene_strand, gene=gene_id) %>% 
		select(gene, chr, start, end, symbol, strand, gene_type)
	test.res<-merge(res, annot.sn0, by="gene") 
	#ick! - only 11195 merged out of 22211 have annotation!  GTEX ensembl ID not the same as our snRNA
	
#  Using non-dotted Ensembl gene IDs since that's what we used to match up our bulk muscle hg38 data
annot.sn<-readRDS("~/sn_muscle_2023/data/annot.sn.Rds") %>% select(-strand)
res1<-res %>% mutate(nodot=word(gene,1,sep=fixed(".")))	
res2<-merge(res1, annot.sn, by.x="nodot", by.y="gene", all.x=T) %>% arrange(pvalue) %>% select(-nodot)
#22109 out of 22211 have annotation -much better

write.table(format(res2, digits=8), 
file=paste0("~/sn_muscle_2023/output/DESeq.RNA/gtex/hg38.filter22211.790gtex.SEX.M.results.tab"), sep="\t", quote=F, col.names=T, row.names=F)


```

